---
title: "Dataset Information"
author: "Micah Fadrigo"
date: "5/9/2022"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(tidyverse)
library(lubridate)
library(dplyr)
```

## Test Positivity Data From PCR Tests

```{r}
# Read in data
# insert path to dataset on local computer 
pcr_data <- read_csv("/Users/micahfadrigo/Desktop/STATS170B_COVID_PROJECT/DATA/All PCR tests updated 1.26.21.csv")

glimpse(pcr_data) # get to know variables
head(pcr_data) # first 6 observations
tail(pcr_data) # last 6 observations
```

__There are `r nrow(pcr_data)` observations and `r ncol(pcr_data)` variables in the dataset. Each observation represents an individual in Orange County who has taken a PCR Test.__

```{r}
pcr_data %>% 
  arrange(ymd(pcr_data$Specimen.Collected.Date)) %>% 
  head()
```

```{r}
pcr_data %>% 
  arrange(ymd(pcr_data$Specimen.Collected.Date)) %>% 
  tail(1)
```

## Mortality Data

```{r}
# Read in data
# insert path to dataset on local computer 
mortality_data <- read_csv("/Users/micahfadrigo/Desktop/STATS170B_COVID_PROJECT/DATA/1.26.21 release to UCI team.csv")

glimpse(mortality_data) # get to know variables
head(mortality_data) # first 6 observations
tail(mortality_data) # last 6 observations
```

__There are `r nrow(mortality_data)` observations and `r ncol(mortality_data)` variables in the dataset. Each observation represents an individual in Orange County who has taken a PCR Test that tested positive for COVID-19.__
