---
title: "Complete analysis: Factors associated with morality among people diagnosed
  with COVID-19"
output: 
  pdf_document:
    extra_dependencies:
      amsmath 
params:
  reclean_data: FALSE
  first_test_date: "2020-03-01"
  last_test_date: "2020-08-16"
  refit_model1: FALSE
  refit_model2: FALSE
  refit_model3: FALSE
  refit_model4: FALSE
  refit_model5: FALSE
  daniel_est_beds: 4879
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  error = FALSE,
  fig.pos = "H"
)
```


```{r loading-libraries}
# data-wrangling-function chunk
library(tidyverse)
library(here)
library(lubridate)
library(data.table)
# missing-data-table chunk
library(knitr)
library(kableExtra)
usepackage_latex("threeparttable")
# fit-model1 chunk
library(sandwich)
library(tictoc)
library(lme4)
library(BRRR)
library(car) # I don't think I need this anymore
library(mgcv)
```


```{r data-wrangling-function}
read_pos_pcr <- function(file_path, start_date, end_date) {
  # Read in data --------------------------------------------------------------------
  pos_results_original <- read_csv(
    file_path,
    col_types = cols(
      .default = col_skip(),
      Age = col_double(),
      Gender = col_character(),
      Ethnicity = col_character(),
      Race = col_character(),
      ReportedCity = col_character(),
      Zip = col_double(),
      SpCollDt = col_date(),
      DeathDueCOVID = col_character(),
      DtDeath = col_date(),
      unique_num = col_double()
    )
  ) 
  
  # Mutate and save desired variables ------------------------------------------------------
  full_num_data_cases <- pos_results_original %>% 
    filter(SpCollDt >= ymd(start_date) & SpCollDt <= ymd(end_date)) %>% 
    nrow()

  
  pos_results_original$Race[is.na(pos_results_original$Race)] <- "Unknown"
  pos_results_original$Ethnicity[is.na(pos_results_original$Ethnicity)] <- "Unknown"

  # All classified as Hispanic or Latino
  hispanic_race_unknown <- (
    (pos_results_original$Race == "Other" & 
     pos_results_original$Ethnicity == "Hispanic or Latino") |
    (pos_results_original$Race == "Unknown" & 
     pos_results_original$Ethnicity == "Hispanic or Latino") |
    (pos_results_original$Race == "Multiple Races" & 
     pos_results_original$Ethnicity == "Hispanic or Latino") 
  )
  
  # All classified as Unknown
  non_hispanic_unknown <- (
    (pos_results_original$Race == "Unknown" & 
     pos_results_original$Ethnicity != "Hispanic or Latino") |
    (pos_results_original$Race == "Multiple Races" & 
     pos_results_original$Ethnicity != "Hispanic or Latino") |
    (pos_results_original$Race == "Other" & 
     pos_results_original$Ethnicity != "Hispanic or Latino")
  )
  
  pos_results_original_new_race <- data.frame(
    pos_results_original, 
    "race1" = str_to_lower(pos_results_original$Race)
  )
  
  pos_results_original_new_race$race1[hispanic_race_unknown] <- "hispanic or latino"
  pos_results_original_new_race$race1[non_hispanic_unknown] <- "unknown"

  pos_results_original_new_race$DeathDueCOVID <- replace_na(
    pos_results_original_new_race$DeathDueCOVID,
    "n"
  )
  pos_results_original_new_race$Age <- replace_na(
    pos_results_original_new_race$Age,
    -999
  )
  pos_results_original_new_race$Zip <- replace_na(
    pos_results_original_new_race$Zip,
    "unknown"
  )
  
  
  
  
  pos_results_adjusted <- pos_results_original_new_race %>% 
    mutate(death_due_to_covid = factor(
      str_to_lower(DeathDueCOVID), 
      levels = c("y", "n")
      )     
    ) %>% 
    mutate(sex = fct_collapse(
      str_to_lower(Gender),
      male = "m",
      female = "f",
      unknown = c("o", "u")
    )) %>%
    mutate(race = factor(
      race1, 
      levels = c(
        "white",
        "asian",
        "black or african american",
        "hispanic or latino",
        "american indian or alaska native",
        "native hawaiian or other pacific islander",
        "unknown"
      ),
    )) %>% 
    mutate(death_date = replace_na(DtDeath, ymd("9999/09/09"))) %>% 
    mutate(           # time_days = days from start date to date they tested positive
      time_days = as.integer(round(difftime(
        SpCollDt, 
        start_date, 
        units = "days"
      )))
    ) %>%
    select(
      unique_num,
      test_date = SpCollDt,
      death_date,
      death_due_to_covid,
      zip = Zip,
      time_days,
      age = Age,
      sex,
      race,
      reported_city = ReportedCity
    ) %>% 
    filter((test_date >= ymd(start_date)) & (test_date <= ymd(end_date))) %>% 
    filter(sex != "unknown") %>% 
    filter(age != -999) %>% 
    filter(zip != "unknown") %>% 
    mutate(zip = str_sub(zip, end = 5)) %>% 
    arrange(test_date) 
  
  
  # Add zip code level data and merge with pcr results ----------------------------------
  zip_area_oc <- read_csv(
    here("data", "zip-area2.csv"),
    col_types = cols(
      .default = col_skip(),
      NAME = col_character(),
      Zip = col_character(),
      AreaKm = col_double()
    )
  ) %>%
    select(name = NAME, zip = Zip, area_km = AreaKm)
  
  zip_pop_oc <- read_csv(
    here("data", "zip-pop.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      Population = col_integer()
    )
  ) %>%
    drop_na() %>%
    mutate(population = Population / 1000) %>%
    select(zip = Zip, population)
  
  zip_data_merged <- merge(x = zip_area_oc, y = zip_pop_oc, by = "zip")
  zip_data_merged$pop_density <- zip_data_merged$population / zip_data_merged$area_km
  
  zip_income_oc <- read_csv(
    here("data", "income-by-zip2.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      IncomeMed = col_integer(),
      IncPeriodofMeas = col_character()
    )
  ) %>%
    mutate(med_income = IncomeMed / 10000) %>%
    filter(IncPeriodofMeas == "2014-2018") %>%
    select(zip = Zip, med_income)
  
  zip_data_merged <- merge(x = zip_data_merged, y = zip_income_oc, by = "zip")

  zip_education_oc <- read_csv(
    here("data", "education-by-zip.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      PercentBach = col_double()
    )
  ) %>%
    select(zip = Zip, percent_bachelors = PercentBach)
  
  zip_data_merged <- merge(x = zip_data_merged, y = zip_education_oc, by = "zip")

  zip_insurance_oc <- read_csv(
    here("data", "insurance-by-zip.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      PercentInsured = col_double()
    )
  ) %>%                           
    select(zip = Zip, percent_insured = PercentInsured)
  
  zip_data_merged <- merge(x = zip_data_merged, y = zip_insurance_oc, by = "zip")

  
  #3 rows in 92678 zipcode we don't have area data for
  pos_results_adjusted$old_zip <- pos_results_adjusted$zip
  pos_results_adjusted$zip[pos_results_adjusted$old_zip == "92678"] <- "92679"
  pos_results_adjusted$zip[pos_results_adjusted$old_zip == "90743"] <- "90742" #############################
  pos_results_merged <- merge(x = pos_results_adjusted, y = zip_data_merged, by = "zip")
  pos_results_merged$old_zip <- factor(pos_results_merged$old_zip)
  pos_results_merged$zip <- factor(pos_results_merged$zip)
  
  
  # Read and filter hospital bed data -----------------------------------------------------------
  # OC data
  hos_bed_gov <- read_csv(
    here("data", "covid19-hospital-data-gov.csv"),
    na = c("", " "),
    col_types = cols(
      .default = col_skip(),
      county = col_character(),
      todays_date = col_date("%Y-%m-%d T %H:%M:%S"),
      hospitalized_covid_patients = col_double(),
      all_hospital_beds = col_double()
    )
  ) %>% 
    filter(county == "Orange") 
  
  first_date <- sort(hos_bed_gov$todays_date)[1]
  if(month(params$first_test_date) == month(first_date) & day(params$first_test_date) < day(first_date)) {
    dates_missing <- seq(as.Date(params$first_test_date), first_date, by = "days")
    dates_missing <- dates_missing[-length(dates_missing)]
    num_dm <- length(dates_missing)
    missing_rows <- data.frame(rep("Orange", num_dm), dates_missing, rep(NA, num_dm), rep(NA, num_dm))
    colnames(missing_rows) <- colnames(hos_bed_gov)
    hos_bed_gov <- rbind(missing_rows, hos_bed_gov)
  } else {
    print("Error: fix hospital data dates")
  }

  # Beds were not recorded before 2020-04-20. Daniel has estimated
  hos_bed_gov <- hos_bed_gov %>% 
    mutate(avail_beds = ifelse(
      is.na(all_hospital_beds - hospitalized_covid_patients), 
      params$daniel_est_beds,
      all_hospital_beds - hospitalized_covid_patients
    )) %>% 
    select(
      test_date = todays_date,
      avail_beds
    )
  
  pos_results_merged <- merge(pos_results_merged, hos_bed_gov, by = "test_date")
  
  
  # Group and scale variables -------------------------------------------------------------------------
  age_breaks <- c(0, 5, 10, 15, 20, 25, 30, 35, 40, 50, 60, 70, 80, 200)
  age_labels <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39",
                  "40-49","50-59","60-69","70-79","80+")
  setDT(pos_results_merged)[, age_group := cut(
    age, 
    breaks = age_breaks, 
    right = FALSE, 
    labels = age_labels
    )]
  
  pos_results_merged$age_group <- factor(
    pos_results_merged$age_group,
    levels = age_labels
    )
  
  pos_results_merged$adj_time_days <- scale(
    pos_results_merged$time_days,
    center = TRUE,
    scale = TRUE
    )
  
  pos_results_merged$adj_pop_density <- scale(
    pos_results_merged$pop_density, 
    center = TRUE, 
    scale = TRUE
    )
  
  pos_results_merged$adj_med_income <- scale(
    pos_results_merged$med_income, 
    center = TRUE, 
    scale = TRUE
    )
  
  pos_results_merged$adj_perc_bach <- scale(
    pos_results_merged$percent_bachelors,
    center = TRUE,
    scale = TRUE
    )
  
  pos_results_merged$adj_perc_bach_quar <- with(
    pos_results_merged,
    cut(adj_perc_bach,
        breaks = quantile(adj_perc_bach, 
        probs = seq(0, 1, by = 0.25)),
        include.lowest = TRUE,
        labels = c("Q1", "Q2", "Q3", "Q4"))
    )
  
  pos_results_merged$adj_perc_insured <- scale(
    pos_results_merged$percent_insured,
    center = TRUE,
    scale = TRUE
    )
  
  pos_results_merged$adj_perc_insured_quar <- with(
    pos_results_merged,
    cut(
      adj_perc_insured,
      breaks = quantile(adj_perc_insured, probs = seq(0, 1, by = 0.25)),
      include.lowest = TRUE,
      labels = c("Q1", "Q2", "Q3", "Q4")
      )
    )
  
  pos_results_merged$adj_age <- scale(pos_results_merged$age, center = TRUE, scale = TRUE)
  pos_results_merged$adj_avail_beds <- scale(pos_results_merged$avail_beds, center = TRUE, scale = TRUE)
  
  
  # Count missing values ------------------------------------------------------------------------------------
  missing_vec <- c(
    full_num_data_cases , # number of observations with test date between start date and end date
    full_num_data_cases - nrow(pos_results_adjusted),  # number of NA for age, sex and zip
    nrow(pos_results_adjusted) - nrow(pos_results_merged) # number of invalid or non Orange County zip codes
  )
  names(missing_vec) <- c(
    "full_num_data_cases", 
    "num_na_cases_removed", 
    "num_bad_zip_cases_removed"
  )
  
  
  list(
    "pos_results_merged" = pos_results_merged, 
    "zip_data_merged" = zip_data_merged, 
    "missing" = missing_vec
  )
}

```


```{r load-clean-data}
if (params$reclean_data) {
  pos_pcr_and_zip <- read_pos_pcr(
    file_path = here("data", "all-positive-tests-update-2020-11-09.csv"),
    start_date = params$first_test_date,
    end_date = params$last_test_date
  )
  save(pos_pcr_and_zip, file = here("data", "cleaned_process_pos_pcr_data.Rdata"))
} else {
  load(file = here("data", "cleaned_process_pos_pcr_data.Rdata"))
  
  pos_pcr <- data.frame(pos_pcr_and_zip[["pos_results_merged"]])
  pos_zip <- data.frame(pos_pcr_and_zip[["zip_data_merged"]])
  pos_missing <- pos_pcr_and_zip[["missing"]]
}
```


```{r missing-data-table}
usable_counts <- c(
  pos_missing[1],
  pos_missing[1] - pos_missing[2],
  pos_missing[1] - sum(pos_missing[2:3])
)

tab_obs <- data.frame(
  c(NA, pos_missing[2:3]), 
  usable_counts
)
rownames(tab_obs) <- c(
  "Unmodified",
  "Missing or Inconclusive[note]",
  "Invalid Zip Code"
)

kable(
  tab_obs,
  col.names = c(
    "With Issue", 
    "After Removal"
  ),
  format = "latex",
  caption = paste(
    "Number of positive COVID-19 tests in Orange County from ",
      params$first_test_date, " to ", params$last_test_date,
      ". Includes number removed from data by type of issue for this analysis.",
    sep = "")
  ) %>%
  kable_styling(latex_options = c("HOLD_position")) %>% 
  add_header_above(c(" " = 1, "Number of Tests" = 2), line = TRUE) %>% 
  column_spec(column = 2:3, width = "5.5cm") %>% 
  add_footnote(
    c(
      "Number without conclusive test result or complete demographic information"
      ),
    notation = "symbol",
    threeparttable = TRUE
    )
```


```{r fit0-results-function} 
compute_ci_logistic <- function(model, alpha = 0.05, param_names, include_pvalues = FALSE) {
  #does not include intercept or median income
  model_summary <- summary(model)
  coeffs <- model_summary$coefficients[-1, 1]
  se_robust <- sqrt(diag(sandwich(model)))[-1]
  model_sum <- data.frame("odds" = exp(coeffs),
                          "lower_bound" = exp(coeffs - qnorm(1 - alpha / 2) * se_robust),
                          "upper_bound" = exp(coeffs + qnorm(1 - alpha / 2) * se_robust),
                          "p-value" =  model_summary$coefficients[-1, 4],
                          "param_names" = param_names)
  if(include_pvalues){
    return("model_sum" = model_sum)
  } else {
    return("model_sum" = model_sum[, -4])
  }
}


```


For simplicity let  
$O_{i}$ be the odds of mortality among individuals who tested positive for COVID-19 in Orange County.  
$\overrightarrow{\beta}_{(3-9)} = (\beta_{\text{Asian}}, \beta_{\text{Black}}, \beta_{\text{Hispanic}}, \beta_{\text{Native American}}, \beta_{\text{Pacific Islander}}, \beta_{\text{Other Race}}, \beta_{\text{Unknown}})$  
$\overrightarrow{\beta}_{(10-12)} = (\beta_{\text{\% with College Degree Quartile2}}, \beta_{\text{\% with College Degree Quartile 3}}, \beta_{\text{\% with College Degree Quartile 4}})$  
$\overrightarrow{\beta}_{(13-15)} = (\beta_{\text{\% with Insurance Quartile 2}}, \beta_{\text{\% with Insurance Quartile 3}}, \beta_{\text{\% with Insurance Quartile 4}})$

Model 0:
\begin{equation}
  \label{eq:Model0}
  \begin{split}
    \log(O_i) &= \beta_0 + \beta_{Age}\text{Age}_i + \beta_{Sex}\text{Sex}_i + \overrightarrow{\beta}_{(3-9)}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{(10-12)}\overrightarrow{\text{\% with College Degree Quartile}}_i + \overrightarrow{\beta}_{(13-15)}\overrightarrow{\text{\% with Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{Time}\text{Time}_i\\
		&+ \beta_{Available Hospital Beds}\text{Available Hospital Beds}_i,
  \end{split}
\end{equation}
without a random intercept for zip code.

```{r fit-model0}
if (params$refit_model1) {
  tic()
  fit0 <- glm(
    formula = death_due_to_covid  ~ adj_age + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days + adj_avail_beds,
    family = binomial, 
    data = pos_pcr
    )
  toc()
  skrrrahh("drummaboy")
  
  save(fit0, file = here("analysis/mortality-regression-results", "fit0.Rdata"))
} else {
  load(file = here("analysis/mortality-regression-results", "fit0.Rdata"))
}

parameter_names <- c(
  "Age",
  "Male",
  "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Unknown",
  "2nd Quartile", "3rd Quartile", "4th Quartile",
  " 2nd Quartile", " 3rd Quartile", " 4th Quartile",
  "Population Density", 
  "Median Income",
  "Days",
  "Available Beds"
  )
summary_fit0 <- compute_ci_logistic(fit0, param_names = parameter_names)

```


Model 1:
\begin{equation}
  \label{eq:Model1}
  \begin{split}
    \log(O_i) &= \beta_0 + \beta_{Age}\text{Age}_i + \beta_{Sex}\text{Sex}_i + \overrightarrow{\beta}_{(3-9)}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{(10-12)}\overrightarrow{\text{\% with College Degree Quartile}}_i + \overrightarrow{\beta}_{(13-15)}\overrightarrow{\text{\% with Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{Time}\text{Time}_i\\
		&+ \beta_{Available Hospital Beds}\text{Available Hospital Beds}_i,
  \end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model1, eval = FALSE}
if (params$refit_model1) {
  tic()
  fit1a <- glmer(
    formula = death_due_to_covid  ~ adj_age + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days + adj_avail_beds +
              (1 | zip),              
    family = binomial, 
    data = pos_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  skrrrahh("drummaboy")
 #  Model failed to converge with max|grad| = 0.078488 (tol = 0.002, component 1)Model is nearly unidentifiable: very large eigenvalue
 # - Rescale variables?
 #  2042.06 sec elapsed
  tic()
  fit1 <- allFit(fit1a)
  toc()
  skrrrahh("drummaboy")
  
  save(fit1, file = here("analysis/mortality-regression-results", "fit1.Rdata"))
} else {
  load(file = here("analysis/mortality-regression-results", "fit1.Rdata"))
}
```


```{r fit0-model-diagnostics, eval = FALSE}
std_res0 <- rstandard(fit0)
qqnorm(std_res0)
plot(fit0, which = 4, id.n = 3)


model.data <- augment(fit0) %>% 
  mutate(index = 1:n()) 

model.data %>% top_n(3, .cooksd)

ggplot(model.data, aes(index, .std.resid)) + 
  geom_point(aes(color = death_due_to_covid), alpha = .5) +
  theme_bw()

model.data %>% 
  filter(abs(.std.resid) > 4)
```


```{r final-model-results-table}
# Not ready yet!
fit_final_model <- summary_fit0
fm_sum <- summary(fit0)

rr_fm <- round(fit_final_model[, "odds"], 3)
ci_lb_fm <- round(fit_final_model[, "lower_bound"], 2)
ci_ub_fm <- round(fit_final_model[, "upper_bound"], 2)
#p_values_fm <- fit_final_model$p.value
#p_values_fm <- ifelse(p_values_fm <= 0.001, "0.000", paste(round(p_values_fm, 3)))

reference_group <- c(
  FALSE,
  TRUE, FALSE,
  TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  FALSE,
  FALSE,
  FALSE,
  FALSE
  )

table_fm <- data.frame(matrix(NA, nrow = nrow(fit_final_model) + sum(reference_group), ncol = 3))
colnames(table_fm) <- c("COVID-19 Deaths", "Total", "with (95% CI[note])")
rownames(table_fm) <- c(
  "Age",
  "Female", "Male",
  "White", "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Unknown",
  "1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile",
  "1st Quartile ", "2nd Quartile ", "3rd Quartile ", "4th Quartile ",
  "Zip Code Pop Density (1000ppl/km^2)",
  "Adjusted Median Income",
  "Adjusted Time (Days)",
  "Available Beds"
  )

          
non_ref_counter <- 1
for(i in 1:nrow(table_fm)) {
  if (reference_group[i]) {
    #table_fm[i, 1] <- "Reference"
    table_fm[i, 3] <- "Reference"
  } else {
    table_fm[i, 3] <- paste(rr_fm[non_ref_counter], " (", 
                          ci_lb_fm[non_ref_counter], ", ", 
                          ci_ub_fm[non_ref_counter], ")", 
                          sep = "")
    #table_fm[i, 2] <- p_values_fm[non_ref_counter]
    non_ref_counter <- non_ref_counter + 1
  }
}      

variable_counts <- data.frame(matrix(NA,  nrow = nrow(fit_final_model) + sum(reference_group), ncol = 2))
dead_pcr <- pos_pcr[pos_pcr$death_due_to_covid == "y", ]
quant_vars <- c("adj_age", "adj_pop_density", "adj_med_income", "adj_time_days", "adj_avail_beds")
order_quant_vars <- c(1, 19:22)
variable_counts[order_quant_vars, c(1, 2)] <- c("", "")

qual_vars <- c("sex", "race", "adj_perc_bach_quar", "adj_perc_insured_quar")
order_qual_vars <- (1:nrow(table_fm))[-order_quant_vars]

qual_vars_ordered <- c(
  "sex.female", "sex.male", 
  "race.white", "race.asian",
  "race.black or african american", "race.hispanic or latino",
  "race.american indian or alaska native", "race.native hawaiian or other pacific islander",
  "race.unknown",
  "adj_perc_bach_quar.Q1", "adj_perc_bach_quar.Q2",
  "adj_perc_bach_quar.Q3", "adj_perc_bach_quar.Q4",
  "adj_perc_insured_quar.Q1", "adj_perc_insured_quar.Q2",
  "adj_perc_insured_quar.Q3", "adj_perc_insured_quar.Q4"
)
 
calc_freq <- function(df, order_vars) {
  qual_count <- unlist(apply(df, MARGIN = 2, FUN = table))

  qual_count <- qual_count[order_vars]
  qual_freq <- round(100 * qual_count / nrow(df), 2)
  paste(qual_count, " (", qual_freq, "%)", sep = "")
}

variable_counts[order_qual_vars, 2] <- calc_freq(pos_pcr[, qual_vars], qual_vars_ordered)
variable_counts[order_qual_vars, 1] <- calc_freq(dead_pcr[, qual_vars], qual_vars_ordered)

table_fm[, 2] <- variable_counts[, 2]
table_fm[, 1] <- variable_counts[, 1]


intercept <- paste0(
  round(exp(fm_sum$coefficients[1, 1]), 3),
  " (",
  round(exp(fm_sum$coefficients[1, 1] - qnorm(0.975) * sqrt(diag(sandwich(fit0)))[1]), 2),
  ",",
  round(exp(fm_sum$coefficients[1, 1] + qnorm(0.975) * sqrt(diag(sandwich(fit0)))[1]), 2),
  ")"
)


kable(
  table_fm, 
  format = "latex", 
  caption = paste(
    "Model 0 regression estimation of adjusted odds ratio of mortality given tested positive for COVID-19 in Orange County.", 
    sep = ""
    )
  ) %>%
  kable_styling(latex_options = c("HOLD_position"), font_size = 11) %>%
  pack_rows("Sex", 2, 3, bold = FALSE) %>%
  pack_rows("Race", 4, 10, bold = FALSE) %>%
  pack_rows("% with College Degree[note]", 11, 14, bold = FALSE) %>%
  pack_rows("% with Insurance[note]", 15, 18, bold = FALSE) %>% 
  add_header_above(c(" " = 1, "Counts" = 2, "Adjusted Odds Ratio[note]" = 1), line = FALSE) %>% 
  add_footnote(
    c(
      paste0(
        "Adjusted for all covariates listed plus zip code estimated median income and time of test in days. Model intercept represents odds of death for a white female diagnosed with SARS-CoV-2 in the 0 to 4 age group in a zip code in the first quartile of college degree and insured with the average population density. The odds of this individual testing dying is estimated to be ",
        intercept,
        collapse = ""
      ), 
      "95% Confidence Interval",
      "Esimated percent of people with a bachelor's degree in an individual's zip code ",
      "Esimated percent of people with medical insurance in an individual's zip code "
      ),
    notation = "symbol"
  )  
```

