---
title: "Factors associated with testing positive for COVID-19 in Orange County"
date: "11/4/2020"
output: 
  pdf_document:
    extra_dependencies:
      amsmath 
params:
  reclean_data: FALSE
  first_test_date: "2020-03-01"
  last_test_date: "2020-08-16"
  refit_model1: FALSE
  refit_model2: FALSE
  refit_model3: FALSE
  refit_model4: FALSE
  refit_model5: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      error = FALSE,
                      fig.pos = "H")
```


```{r loading-libraries}
# data-wrangling-function chunk
library(tidyverse)
library(here)
library(lubridate)
library(data.table)
# missing-data-table chunk
library(knitr)
library(kableExtra)
# fit-model1 chunk
library(tictoc)
library(lme4)
library(BRRR)
library(car) # I dont think I need this anymore
library(mgcv)

library(ggplot2)
```

```{r data-wrangling-function}
# Test results need to be categorized ------------------------------------------------------
negative_test_synonyms <- c(
  "not detected",
  "negative",
  "coronavirus 2019 novel not detected",
  "negative (qualifier value)",
  "not detected (qualifier value)",
  "sars cov-2 negative",
  "undetected",
  "inst_negative",
  "neg-see report",
  "sars-cov-2 rna not detected by naa",
  "none detected", 
  "not detected in pooled specimen"
  )

positive_test_synonyms <- c(
  "detected",
  "coronavirus 2019 novel positive",
  "positive",
  "positive (qualifier value)",
  "sars cov-2 positive",
  "detected (qualifier value)",
  "presumptive pos",
  "positive for 2019-ncov",
  "presumptive positive",
  "coronavirus 2019 novel presumptive pos",
  "coronavirus 2019 novel detected",
  "yes",
  "coronavirus 2019 novel",
  "presumptive positive for 2019-ncov",
  "sars cov-2 presumptive pos",
  "presumptive pos. for 2019-ncov",
  "presumptive positive (qualifier value)",
  "presumptive detected",
  "reactive",
  "sars-cov-2"
  )

other_test_synonyms <- c(
  "inconclusive",
  "indeterminate",
  "specimen unsatisfactory",
  "invalid",
  "test not performed",
  "not provided (qualifier value)",
  "see comment",
  "tnp",
  "coronavirus 2019 novel inconclusive",
  "not tested",
  "phoned results (and readback confirmed) to:",
  "see note",
  "clotted",
  "coronavirus 2019 novel unsatisfactory",
  "equivocal",
  "non reactive",
  "result comments",
  "sars cov-2 inconclusive",
  "test not done",
  "test not perf",
  "not pregnant",
  "biofiresarsneg",
  "equivocal result",
  "coronavirus 2019 novel inconcluside",
  "unsatisfactory",
  "undefined",
  "*corrected report* by",
  "specimen unsatifactory for evaluation",
  "warning....please disregard results.",
  "presumptive result to be confirmed",
  "indeterminate (qualifier value)", 
  "invalid result", 
  "specimen unsatisfactory for evaluation"
  )


read_all_pcr <- function(file_path, start_date, end_date) {
  pcr_results_original <- read_csv(
    file_path,
    col_types = cols(
      .default = col_skip(),
      PersonId = col_character(),
      Age = col_integer(),
      Sex = col_character(),
      Ethnicity = col_character(),
      Race = col_character(),
      Specimen.Collected.Date = col_date("%m-%d-%Y"),
      Resulted.Organism = col_character(),
      Zip = col_character()
      )
    ) 
  
  full_num_data_cases <- pcr_results_original %>% 
    filter(Specimen.Collected.Date >= ymd(start_date) & Specimen.Collected.Date <= ymd(end_date)) %>% 
    nrow()
  
  pcr_results_original$Race[is.na(pcr_results_original$Race)] <- "Unknown"
  pcr_results_original$Ethnicity[is.na(pcr_results_original$Ethnicity)] <- "Unknown"

  hispanic_race_unknown <- (
    (pcr_results_original$Race == "Other" & 
     pcr_results_original$Ethnicity == "Hispanic or Latino") |
    (pcr_results_original$Race == "Unknown" & 
     pcr_results_original$Ethnicity == "Hispanic or Latino") |
    (pcr_results_original$Race == "Multiple Races" & 
     pcr_results_original$Ethnicity == "Hispanic or Latino") 
    )
  
  non_hispanic_unknown <- (
    (pcr_results_original$Race == "Unknown" & 
     pcr_results_original$Ethnicity != "Hispanic or Latino") |
    (pcr_results_original$Race == "Multiple Races" & 
     pcr_results_original$Ethnicity != "Hispanic or Latino")
    )
  
  pcr_results_original_new_race <- data.frame(
    pcr_results_original, 
    "race1" = str_to_lower(pcr_results_original$Race)
    )
  pcr_results_original_new_race$race1[hispanic_race_unknown] <- "hispanic or latino"
  pcr_results_original_new_race$race1[non_hispanic_unknown] <- "unknown"
  
  pcr_results_adjusted <- pcr_results_original_new_race %>%
    filter(!is.na(Resulted.Organism)) %>%
    mutate(test_result = fct_collapse(
      str_to_lower(Resulted.Organism),
      negative = negative_test_synonyms,
      positive = positive_test_synonyms,
      unknown = other_test_synonyms) 
      ) %>%
    mutate(sex = fct_collapse(
      str_to_lower(Sex),
      male = "m",
      female = "f",
      unknown = c("d", "g", "i", "tf", "tm", "u")
      )) %>%
    mutate(race = factor(
      race1, 
      levels = c(
        "white",
        "asian",
        "black or african american",
        "hispanic or latino",
        "american indian or alaska native",
        "native hawaiian or other pacific islander",
        "other",
        "unknown"
        ),
      )) %>% 
    mutate(
      time_days = as.integer(round(difftime(
        Specimen.Collected.Date, 
        start_date, 
        units = "days"
        )))
      ) %>%
    select(
      id = PersonId, 
      posted_date = Specimen.Collected.Date, 
      time_days,
      test_result, 
      age = Age,
      sex,
      race,
      ethnicity = Ethnicity,
      zip = Zip,
      ) %>%
    filter((posted_date >= ymd(start_date)) & (posted_date <= ymd(end_date))) %>%
    filter(test_result != "unknown") %>%
    filter(sex != "unknown") %>%
    mutate(zip = str_sub(zip, end = 5)) %>%
    drop_na() %>% 
    group_by(id) %>%
    arrange(posted_date) %>%
    ungroup()
  
  
  if(length(levels(pcr_results_adjusted$test_result)) != 3) warning("New test result category not accounted for.")

  pcr_results_adjusted$covid_positive <- ifelse( pcr_results_adjusted$test_result == "positive", 1, 0)
  
  
  # Extract df of all observations with an id that has inconsistencies for the demographic variables (age, sex, race)
  pcr_rep_id <- pcr_results_adjusted[(
    duplicated(pcr_results_adjusted$id, fromLast=TRUE) | 
    duplicated(pcr_results_adjusted$id)
    ), ]

  pcr_rep_id2 <- pcr_rep_id %>%
    group_by(id) %>% 
    mutate(reasonable_ages = diff(range(age)) <= 1) %>%
    mutate(identical_race = n_distinct(race) == 1) %>%
    mutate(identical_sex = n_distinct(sex) == 1) %>%
    ungroup()
  
  pcr_inconsistent <- data.frame(
    pcr_rep_id2[(
      !pcr_rep_id2$reasonable_ages | 
      !pcr_rep_id2$identical_sex | 
      !pcr_rep_id2$identical_race
      ), ]
    )
  
  inconsistent_counts <- c(
    length(unique(pcr_inconsistent$id[(!pcr_inconsistent$reasonable_ages)])),
    length(unique(pcr_inconsistent$id[(!pcr_inconsistent$identical_sex)])),
    length(unique(pcr_inconsistent$id[(!pcr_inconsistent$identical_race)]))
    )
  names(inconsistent_counts) <- c(
    "number_id_age_inconsistencies", 
    "number_id_sex_inconsistencies",
    "number_id_race_inconsistencies"
    )
  
  pcr_results_consistent <- pcr_results_adjusted[!(pcr_results_adjusted$id %in% pcr_inconsistent$id), ]

  
  # Need to keep all observations for a person up to and including their first positive
  first_pos <- pcr_results_consistent %>%
    filter(test_result == "positive") %>%
    group_by(id) %>%
    summarise(first_pos = min(posted_date))
  
  pcr_results_reduced <- left_join(pcr_results_adjusted, first_pos) %>%
    mutate(first_pos = replace_na(first_pos, lubridate::ymd("9999-12-31"))) %>%
    filter(posted_date <= first_pos) %>%
    select(-first_pos) %>%
    distinct()
  
  
  # Add zip code level data and merge with pcr results
  zip_area_oc <- read_csv(
    here("data", "zip-area2.csv"),
    col_types = cols(
      .default = col_skip(),
      NAME = col_character(),
      Zip = col_character(),
      AreaKm = col_double()
      )
    ) %>%
    select(name = NAME, zip = Zip, area_km = AreaKm)
  
  zip_pop_oc <- read_csv(
    here("data", "zip-pop.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      Population = col_integer()
      )
    ) %>%
    drop_na() %>%
    mutate(population = Population / 1000) %>%
    select(zip = Zip, population)
  
  zip_data_merged <- merge(x = zip_area_oc, y = zip_pop_oc, by = "zip")
  zip_data_merged$pop_density <- zip_data_merged$population / zip_data_merged$area_km
  
  zip_income_oc <- read_csv(
    here("data", "income-by-zip2.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      IncomeMed = col_integer(),
      IncPeriodofMeas = col_character()
      )
    ) %>%
    mutate(med_income = IncomeMed / 10000) %>%
    filter(IncPeriodofMeas == "2014-2018") %>%
    select(zip = Zip, med_income)
  
  zip_data_merged <- merge(x = zip_data_merged, y = zip_income_oc, by = "zip")

  zip_education_oc <- read_csv(
    here("data", "education-by-zip.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      PercentBach = col_double()
      )
    ) %>%
    select(zip = Zip, percent_bachelors = PercentBach)
  
  zip_data_merged <- merge(x = zip_data_merged, y = zip_education_oc, by = "zip")

  zip_insurance_oc <- read_csv(
    here("data", "insurance-by-zip.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      PercentInsured = col_double()
      )
    ) %>%                           
    select(zip = Zip, percent_insured = PercentInsured)
  
  zip_data_merged <- merge(x = zip_data_merged, y = zip_insurance_oc, by = "zip")

  
  #23 rows in 92678 zipcode we don't have area data for
  pcr_results_reduced$old_zip <- pcr_results_reduced$zip
  pcr_results_reduced$zip[pcr_results_reduced$old_zip == "92678"] <- "92679"
  pcr_results_merged <- merge(x = pcr_results_reduced, y = zip_data_merged, by = "zip")
  pcr_results_merged$old_zip <- factor(pcr_results_merged$old_zip)
  pcr_results_merged$zip <- factor(pcr_results_merged$zip)
  
  
  missing_counts <- c(
    full_num_data_cases , 
    full_num_data_cases - nrow(pcr_results_adjusted), 
    nrow(pcr_results_adjusted) - nrow(pcr_results_consistent),
    nrow(pcr_results_consistent) - nrow(pcr_results_reduced),
    nrow(pcr_results_reduced) - nrow(pcr_results_merged)
    )
  names(missing_counts) <- c(
    "full_num_data_cases", 
    "num_na_cases_removed", 
    "num_inconsistent_cases_removed",
    "num_consecutive_cases_removed",
    "num_bad_zip_cases_removed"
    )
  
  
  # Group and scale variables 
  age_breaks <- c(0, 5, 10, 15, 20, 25, 30, 35, 40, 50, 60, 70, 80, 200)
  age_labels <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39",
                  "40-49","50-59","60-69","70-79","80+")
  setDT(pcr_results_merged)[, age_group := cut(
    age, 
    breaks = age_breaks, 
    right = FALSE, 
    labels = age_labels
    )]
  
  pcr_results_merged$age_group <- factor(
    pcr_results_merged$age_group,
    levels = age_labels
    )
  
  pcr_results_merged$adj_time_days <- scale(
    pcr_results_merged$time_days,
    center = TRUE,
    scale = TRUE
    )
  
  pcr_results_merged$adj_pop_density <- scale(
    pcr_results_merged$pop_density, 
    center = TRUE, 
    scale = TRUE
    )
  
  pcr_results_merged$adj_med_income <- scale(
    pcr_results_merged$med_income, 
    center = TRUE, 
    scale = TRUE
    )
  
  pcr_results_merged$adj_perc_bach <- scale(
    pcr_results_merged$percent_bachelors,
    center = TRUE,
    scale = TRUE
    )
  
  pcr_results_merged$adj_perc_bach_quar <- with(
    pcr_results_merged,
    cut(adj_perc_bach,
        breaks = quantile(adj_perc_bach, 
        probs = seq(0, 1, by = 0.25)),
        include.lowest = TRUE,
        labels = c("Q1", "Q2", "Q3", "Q4"))
    )
  
  pcr_results_merged$adj_perc_insured <- scale(
    pcr_results_merged$percent_insured,
    center = TRUE,
    scale = TRUE
    )
  
  pcr_results_merged$adj_perc_insured_quar <- with(
    pcr_results_merged,
    cut(
      adj_perc_insured,
      breaks = quantile(adj_perc_insured, probs = seq(0, 1, by = 0.25)),
      include.lowest = TRUE,
      labels = c("Q1", "Q2", "Q3", "Q4")
      )
    )
  

  
  list(
    "pcr_results_merged" = pcr_results_merged, 
    "zip_data_merged" = zip_data_merged, 
    "counts" = missing_counts,
    "inconsistencies" = inconsistent_counts
    )
}
```


```{r load-clean-data}
if (params$reclean_data) {
  all_pcr_and_zip <- read_all_pcr(
    file_path = here("data", "all-elr-pcr-tests-updated-2020-10-05.csv"),
    start_date = params$first_test_date,
    end_date = params$last_test_date
    )
  
  all_pcr <- data.frame(all_pcr_and_zip[["pcr_results_merged"]])
  all_zip <- data.frame(all_pcr_and_zip[["zip_data_merged"]])
  all_counts <- all_pcr_and_zip[["counts"]]
  all_inconsist_ids <- all_pcr_and_zip[["inconsistencies"]]
  save(all_pcr_and_zip, file = here("data", "cleaned_process_pcr_data.Rdata"))
} else {
  load(file = here("data", "cleaned_process_pcr_data.Rdata"))
  
  all_pcr <- data.frame(all_pcr_and_zip[["pcr_results_merged"]])
  all_zip <- data.frame(all_pcr_and_zip[["zip_data_merged"]])
  all_counts <- all_pcr_and_zip[["counts"]]
  all_inconsist_ids <- all_pcr_and_zip[["inconsistencies"]]
}
```


```{r missing-data-table}
tab_obs <- data.frame(matrix( data = NA, ncol = 2, nrow = 5,))
rownames(tab_obs) <- c(
  "Unmodified",
  "Nonresponse[note]",
  "Inconsistent",
  "Consecutive testing[note]",
  "Invalid zip code"
  )
tab_obs[1, 1] <- all_counts["full_num_data_cases"]
tab_obs[1, 2] <- NA
tab_obs[2:5, 1] <- all_counts[-1]
tab_obs[2, 2] <- tab_obs[1, 1] - tab_obs[2, 1]
tab_obs[3, 2] <- tab_obs[2, 2] - tab_obs[3, 1]
tab_obs[4, 2] <- tab_obs[3, 2] - tab_obs[4, 1]
tab_obs[5, 2] <- tab_obs[4, 2] - tab_obs[5, 1]

kable(
  tab_obs,
  col.names = c("Number of tests", "Number of usable tests after removal"),
  format = "latex",
  caption = paste(
    "Number of COVID-19 tests in Orange County from ", 
      params$first_test_date, " to ", params$last_test_date, 
      ". Includes number of removed tests from data by type of issue for this analysis.", 
    sep = "")
  ) %>%
  kable_styling(latex_options = c("HOLD_position")) %>% 
  add_footnote(
    c(
      "Number of tests without test result or complete demographic information",
      "Kept an indivdual's tests up to and including their first positive"
      ),
    notation = "symbol"
    )
```


```{r compute-ci-gam-logistic-function}
param_names <- c(
  "Intercept",
  "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", 
  "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Male",
  "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Other Race", "Unknown Race",
  "Bachelors Q2", "Bachelors Q3", "Bachelors Q4",
  "Insured Q2", "Insured Q3", "Insured Q4",
  "Pop Density"
  )

compute_ci_gam_logistic <- function(model, alpha = 0.05, plot_estimates = TRUE) {
  model_summary <- summary(model)
  in_med_inc <- length(model_summary$p.coeff)
  coeffs <- model_summary$p.coeff[-in_med_inc]
  se <- sqrt(diag(vcov(model, unconditional = TRUE)))[1:length(coeffs)]
  model_sum <- data.frame("odds" = exp(coeffs),
                          "lower_bound" = exp(coeffs - qnorm(1 - alpha / 2) * se),
                          "upper_bound" = exp(coeffs + qnorm(1 - alpha / 2) * se),
                          "p-value" =  model_summary$p.pv[-in_med_inc],
                          "param_names" = param_names)
  if (plot_estimates) {
    ci_plot <- ggplot(model_sum, 
                        aes(x = param_names, y = odds),
                        size = 5, colour = "red") +
    ylim(min(model_sum$lower_bound), max(model_sum$upper_bound)) +
    geom_point() +
    geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), colour = "black") +
    geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
    scale_x_discrete(limits = model_sum$param_names) +
    theme(axis.text.x = element_text(face = "bold", size = 12, angle = 90))
    
    return(list("model_sum" = model_sum, "ci_plot" = ci_plot))
  }
  
  return("model_sum" = model_sum)
}
```

For simplicity let  
$p_{i}$ being the probability of testing positive for COVID-19 in Orange County  
$\overrightarrow{\beta}^{T}_{(1-12)} = (\beta_{\text{Age[5-9]}}, \beta_{\text{Age[10-14]}}, \beta_{\text{Age[15-19]}}, \beta_{\text{Age[20-24]}}, \beta_{\text{Age[25-29]}}, \beta_{\text{Age[30-34]}}, \beta_{\text{Age[35-39]}}\\ \beta_{\text{Age[40-49]}},  \beta_{\text{Age[50-59]}}, \beta_{\text{Age[60-69]}}, \beta_{\text{Age[70-79]}}, \beta_{\text{Age[80+]}})$  
$\overrightarrow{\beta}_{(14-20)} = (\beta_{\text{Asian}}, \beta_{\text{Black}}, \beta_{\text{Hispanic}}, \beta_{\text{NativeAmerican}}, \beta_{\text{PacificIslander}}, \beta_{\text{OtherRace}}, \beta_{\text{Unknown}})$  
$\overrightarrow{\beta}_{(21-23)} = (\beta_{\text{EducationQ2}}, \beta_{\text{EducationQ3}}, \beta_{\text{EducationQ4}})$  
$\overrightarrow{\beta}_{(24-26)} = (\beta_{\text{InsuredQ2}}, \beta_{\text{InsuredQ3}}, \beta_{\text{InsuredQ4}})$


Model 1:
\begin{equation}
  \label{eq:Model1}
  \begin{split}
    \log(\frac{p_i}{1 + p_i}) &= \beta_0 + \overrightarrow{\beta}^{T}_{(1-12)}\overrightarrow{\text{Age Group}}_i + \overrightarrow{\beta}_{13}\text{Male}_i + \overrightarrow{\beta}_{(14-20)}\overrightarrow{\text{Race}}_i\\
    &+ \overrightarrow{\beta}_{(21-23)}\overrightarrow{\text{Education Quartile}}_i + \overrightarrow{\beta}_{(24-26)}\overrightarrow{\text{Insured}}_i + \beta_{27}\text{PopDensity}_i\\
    &+ \beta_{28}\text{MedianIncome}_i\\ 
    &+ \beta_{29}\text{Days}_i,
  \end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model1}
if (params$refit_model1) {
  tic()
  fit_time_lin1 <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days +
              (1 | zip),              
    family = binomial, 
    data = all_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  # 4423.84 sec elapsed
  # Model failed to converge with max|grad| = 0.00298652 (tol = 0.002, component 1)
  tic()
  ss <- getME(fit_time_lin1, c("theta","fixef"))
  fit_time_lin2 <- update(
    fit_time_lin1,
    start = ss,
    control = glmerControl(
      optimizer = "bobyqa",
      optCtrl = list(maxfun = 2e6)
    )
  )
  toc()
  fit_time_lin <- fit_time_lin2
  skrrrahh("drummaboy")
  # 4372.56 sec elapsed
  # updated and saved 11-3-2020
  save(fit_time_lin, file = here("analysis/testing-positive-regression-results", "fit_time_lin.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_lin.Rdata"))
}
```

Model 2:
\begin{equation}
  \label{eq:Model2}
  \begin{split}
    \log(\frac{p_i}{1 + p_i}) &= \beta_0 + \overrightarrow{\beta}^{T}_{(1-12)}\overrightarrow{\text{Age Group}}_i + \overrightarrow{\beta}_{13}\text{Male}_i + \overrightarrow{\beta}_{(14-20)}\overrightarrow{\text{Race}}_i\\
    &+ \overrightarrow{\beta}_{(21-23)}\overrightarrow{\text{Education Quartile}}_i + \overrightarrow{\beta}_{(24-26)}\overrightarrow{\text{Insured}}_i + \beta_{27}\text{PopDensity}_i\\
    &+ \beta_{28}\text{MedianIncome}_i\\ 
    &+ \beta_{29}\text{Days}_i + \beta_{30}(\text{Days}_i)^2,
  \end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model2}
if (params$refit_model2) {
  tic()
  fit_time_quad <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              I(adj_time_days) + I(adj_time_days^2) + 
              (1 | zip),              
    family = binomial, 
    data = all_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  skrrrahh("drummaboy")
  # 4846.83 sec elapsed
  # updated and saved 11-2-2020
  save(fit_time_quad, file = here("analysis/testing-positive-regression-results", "fit_time_quad.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_quad.Rdata"))
}
```

Model 3:
\begin{equation}
  \label{eq:Model3}
  \begin{split}
    \log(\frac{p_i}{1 + p_i}) &= \beta_0 + \overrightarrow{\beta}^{T}_{(1-12)}\overrightarrow{\text{Age Group}}_i + \overrightarrow{\beta}_{13}\text{Male}_i + \overrightarrow{\beta}_{(14-20)}\overrightarrow{\text{Race}}_i\\
    &+ \overrightarrow{\beta}_{(21-23)}\overrightarrow{\text{Education Quartile}}_i + \overrightarrow{\beta}_{(24-26)}\overrightarrow{\text{Insured}}_i + \beta_{27}\text{PopDensity}_i\\
    &+ \beta_{28}\text{MedianIncome}_i\\
    &+ \beta_{29}\text{Days}_i + \beta_{30}(\text{Days}_i)^2\\ 
    &+ \beta_{31}\text{MedianIncome}_i\times\text{Days}_i + \beta_{32}\text{MedianIncome}_i\times(\text{Days}_i)^2,
  \end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model3}
if (params$refit_model3) {
  tic()
  fit_time_quad_inter1 <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days + I(adj_time_days^2) + 
              adj_time_days:adj_med_income + I(adj_time_days^2):adj_med_income +
              (1 | zip),              
    family = binomial, 
    data = all_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  # 5212.67 sec elapsed
  # Model failed to converge with max|grad| = 0.00348316 (tol = 0.002, component 1)
    tic()
  ss <- getME(fit_time_quad_inter1, c("theta","fixef"))
  fit_time_quad_inter2 <- update(
    fit_time_quad_inter1,
    start = ss,
    control = glmerControl(
      optimizer = "bobyqa",
      optCtrl = list(maxfun = 2e6)
    )
  )
  toc()

  fit_time_quad_inter <- fit_time_quad_inter2
  # 3284.9 sec elapsed
  # Model failed to converge with max|grad| = 0.00348316 (tol = 0.002, component 1)
  skrrrahh("drummaboy")
  # updated 11/3/2020
  save(fit_time_quad_inter, file = here("analysis/testing-positive-regression-results", "fit_time_quad_inter.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_quad_inter.Rdata"))
}
```

Model 4:  
Generalized additive model
\begin{equation}
  \label{eq:Model4}
  \begin{split}
    \log(\frac{p_i}{1 + p_i}) &= \beta_0 + \overrightarrow{\beta}^{T}_{(1-12)}\overrightarrow{\text{Age Group}}_i + \overrightarrow{\beta}_{13}\text{Male}_i + \overrightarrow{\beta}_{(14-20)}\overrightarrow{\text{Race}}_i\\
    &+ \overrightarrow{\beta}_{(21-23)}\overrightarrow{\text{Education Quartile}}_i + \overrightarrow{\beta}_{(24-26)}\overrightarrow{\text{Insured}}_i + \beta_{27}\text{PopDensity}_i\\
    &+ \beta_{28}\text{MedianIncome}_i\\ 
    &+ \beta_{29}f(\text{Days}_i),
  \end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model4}
if (params$refit_model4) {
  tic()
  fit_time_gam <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density  + adj_med_income +
              s(time_days, bs = "ts", k = -1) + 
              s(zip, bs = "re"),              
    family = binomial, 
    data = all_pcr,
    method = "REML",
    gamma = 1.5
    )
  toc()
  skrrrahh("drummaboy")
  # 1289.41 sec elapsed
  # updated and saved 11-2-2020
  save(fit_time_gam, file = here("analysis/testing-positive-regression-results", "fit_time_gam.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_gam.Rdata"))
}
```

Model 5:  
Generalized additive model
\begin{equation}
  \label{eq:Model5}
  \begin{split}
    \log(\frac{p_i}{1 + p_i}) &= \beta_0 + \overrightarrow{\beta}^{T}_{(1-12)}\overrightarrow{\text{Age Group}}_i + \overrightarrow{\beta}_{13}\text{Male}_i + \overrightarrow{\beta}_{(14-20)}\overrightarrow{\text{Race}}_i\\
    &+ \overrightarrow{\beta}_{(21-23)}\overrightarrow{\text{Education Quartile}}_i + \overrightarrow{\beta}_{(24-26)}\overrightarrow{\text{Insured}}_i + \beta_{27}\text{PopDensity}_i\\
    &+ \beta_{28}\text{MedianIncome}_i\\ 
    &+ \beta_{29}f(\text{Days}_i)\\
    &+ \beta_{30}f(\text{Days}_i\times\text{MedianIncome}_i),
  \end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model5, results="hide"}
if (params$refit_model5) {
  tic()
  fit_time_gam_inter <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density  + adj_med_income +
              s(time_days) +
              ti(time_days, adj_med_income, bs = "ts") + 
              s(zip, bs = "re"),              
    family = binomial, 
    data = all_pcr,
    method = "REML",
    gamma = 1.5
    )
  toc()
  skrrrahh("drummaboy")
# 4714.62 sec elapsed
# updated and saved 11-2-2020
  save(fit_time_gam_inter, file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter.Rdata"))
}
 
fit_time_gam_inter_info <- compute_ci_gam_logistic(
  model = fit_time_gam_inter
  )
fit_time_gam_inter_summary <- fit_time_gam_inter_info[["model_sum"]]
fit_time_gam_inter_plot <- fit_time_gam_inter_info[["ci_plot"]]
gam_smooth_var_ci <- gam.vcomp(fit_time_gam_inter)
```


```{r bic-table}
compare_bic <- BIC(fit_time_lin, fit_time_quad, fit_time_quad_inter, fit_time_gam, fit_time_gam_inter) 
row.names(compare_bic) <- c("Model 1", "Model 2", "Model 3", "Model 4", "Model 5")

kable(compare_bic,
      col.names = c("Degrees of Freedom", "BIC"),
      format = "latex",
      caption = "Model comparison using BIC supports Model 5 to model odds of testing positive for COVID-19 in Orange County.") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r final-model-results-table}
fit_final_model <- fit_time_gam_inter_summary[-1, ]

rr_fm <- round(fit_final_model[, "odds"], 3)
ci_lb_fm <- round(fit_final_model[, "lower_bound"], 2)
ci_ub_fm <- round(fit_final_model[, "upper_bound"], 2)
#p_values_fm <- fit_final_model$p.value
#p_values_fm <- ifelse(p_values_fm <= 0.001, "0.000", paste(round(p_values_fm, 3)))
# The 5 accounts for the reference groups
table_fm <- data.frame(matrix(NA, nrow = nrow(fit_final_model) + 5, ncol = 3))
colnames(table_fm) <- c("COVID19+", "Total", "with (95% CI[note])")
rownames(table_fm) <- c(
  "0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", 
  "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Female", "Male",
  "White", "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Other", "Unknown",
  "1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile",
  "1st Quartile ", "2nd Quartile ", "3rd Quartile ", "4th Quartile ",
  "Zip Code Pop Density (1000ppl/km^2)"
  )
reference_group <- c(
  TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,
  TRUE, FALSE,
  TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  FALSE
  )
          
non_ref_counter <- 1
for(i in 1:nrow(table_fm)) {
  if (reference_group[i]) {
    #table_fm[i, 1] <- "Reference"
    table_fm[i, 3] <- "Reference"
  } else {
    table_fm[i, 3] <- paste(rr_fm[non_ref_counter], " (", 
                          ci_lb_fm[non_ref_counter], ", ", 
                          ci_ub_fm[non_ref_counter], ")", 
                          sep = "")
    #table_fm[i, 2] <- p_values_fm[non_ref_counter]
    non_ref_counter <- non_ref_counter + 1
  }
}      

variable_counts <- data.frame(matrix(NA,  nrow = nrow(fit_final_model) + 5, ncol = 2))
cov_pos_pcr <- all_pcr[all_pcr$covid_positive == 1, ]
quant_vars <- c("pop_density")
order_quant_vars <- c(nrow(table_fm))
variable_counts[order_quant_vars, c(1, 2)] <- c("", "")

qual_vars <- c("age_group", "sex", "race", "adj_perc_bach_quar", "adj_perc_insured_quar")
order_qual_vars <- (1:nrow(table_fm))[-order_quant_vars]

qual_vars_ordered <- c(
  "age_group.0-4","age_group.5-9", "age_group.10-14", "age_group.15-19",
  "age_group.20-24", "age_group.25-29", "age_group.30-34", "age_group.35-39",
  "age_group.40-49", "age_group.50-59", "age_group.60-69", "age_group.70-79",
  "age_group.80+", 
  "sex.female", "sex.male", 
  "race.white", "race.asian",
  "race.black or african american", "race.hispanic or latino",
  "race.american indian or alaska native", "race.native hawaiian or other pacific islander",
  "race.other", "race.unknown",
  "adj_perc_bach_quar.Q1", "adj_perc_bach_quar.Q2",
  "adj_perc_bach_quar.Q3", "adj_perc_bach_quar.Q4",
  "adj_perc_insured_quar.Q1", "adj_perc_insured_quar.Q2",
  "adj_perc_insured_quar.Q3", "adj_perc_insured_quar.Q4"
  )
 
calc_freq <- function(df, order_vars) {
  qual_count <- unlist(apply(df, MARGIN = 2, FUN = table))

  qual_count <- qual_count[order_vars]
  qual_freq <- round(100 * qual_count / nrow(df), 2)
  paste(qual_count, " (", qual_freq, "%)", sep = "")
}

variable_counts[order_qual_vars, 2] <- calc_freq(all_pcr[, qual_vars], qual_vars_ordered)
variable_counts[order_qual_vars, 1] <- calc_freq(cov_pos_pcr[, qual_vars], qual_vars_ordered)

table_fm[, 2] <- variable_counts[, 2]
table_fm[, 1] <- variable_counts[, 1]


kable(
  table_fm, 
  format = "latex", 
  caption = paste(
    "Model 5 regression estimation of fixed effects for of odds of testing positive for COVID-19 in Orange County. Model includes a smooth term for time and an interaction between median income and time.", 
    sep = ""
    )
  ) %>%
  kable_styling(latex_options = c("HOLD_position"), font_size = 11) %>%
  pack_rows("Age", 1, 13, bold = FALSE) %>%
  pack_rows("Sex", 14,15, bold = FALSE) %>%
  pack_rows("Race", 16, 23, bold = FALSE) %>%
  pack_rows("Zip Code POP with Bachelors", 24, 27, bold = FALSE) %>%
  pack_rows("Zip Code POP with Medical Insurance", 28, 31, bold = FALSE) %>% 
  add_header_above(c(" " = 1, "Counts" = 2, "Adjusted Odds Ratio[note]" = 1), line = FALSE) %>% 
  add_footnote(
    c(
      "Adjusted for all covariates listed plus zip code median income and time of test in days", 
      "Abreviations: CI = confidence interval, POP = percentage of population"
      ),
    notation = "symbol"
  )  
```


```{r final-model-results-plot, fig.align="center", fig.cap="Model 5 regression results for linear fixed effects. This plot excludes the coefficient for median income due to the interaction between median income and time."}
fit_time_gam_inter_plot +
  ylab("Odds of testing COVID positive") +
  xlab("")
```


```{r interaction-plots, fig.align="center", fig.width=7, fig.height=7, fig.cap="Three dimentional visualization of interaction between time (days) and adjusted median income (sample sd) from logistic model 5. COVID-19 test results in Orange County from March 1st to August 16th. Large difference visible in early March between individuals in zip codes with relatively high median incomes and those in zip codes with relatively small median incomes; this difference diminishes as time increase."}
plot(fit_time_gam_inter, scheme = 1, select = 2, 
     xlab = "Time (days)",
     ylab = "Adjusted median income",
     main = "Interaction") 
```



