---
title: "Factors associated with mortality among individuals who tested positive for SARS-CoV-2 in Orange County"
output: 
  pdf_document:
    extra_dependencies:
      amsmath 
params:
  reclean_data: FALSE
  first_test_date: "2020-03-01"
  last_test_date: "2020-08-16"
  refit_model0: FALSE
  refit_model1: FALSE
  daniel_est_beds: 4879
  oc_icu_avail_beds_earliest_val: 131
  oc_hos_covid_pateients_earliest_val: 308
  oc_all_hos_bed_earliest_val: 4213
  oc_icu_full_beds_earliest_val: 71
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  error = FALSE,
  fig.pos = "H"
)
```


```{r loading-libraries}
# data-wrangling-function chunk
library(tidyverse)
library(here)
library(lubridate)
library(data.table)
# missing-data-table chunk
library(knitr)
library(kableExtra)
usepackage_latex("threeparttable")
# fit-model1 chunk
library(sandwich)
library(tictoc)
library(lme4)
#library(BRRR)
library(car) # I don't think I need this anymore
library(mgcv)
```


```{r data-wrangling-function}
read_pos_pcr <- function(file_path, start_date, end_date) {
  # Read in data --------------------------------------------------------------------
  mortality_og <- read_csv(
    file_path,
    col_types = cols(
      .default = col_skip(),
      Age = col_double(),
      Gender = col_character(),
      Ethnicity = col_character(),
      Race = col_character(),
      ReportedCity = col_character(),
      Zip = col_double(),
      SpCollDt = col_date(),
      DeathDueCOVID = col_character(),
      DtDeath = col_date(),
      unique_num = col_character()
    )
  ) %>% 
    mutate(           # time_days = days from start date to date they tested positive
      time_days = as.integer(round(difftime(
        SpCollDt, 
        start_date, 
        units = "days"
      )))
    ) %>% 
    mutate(death_date = replace_na(DtDeath, ymd("9999/09/09"))) %>%
    mutate(decades_old = Age / 10) %>% 
    mutate(DeathDueCOVID = ifelse(is.na(DeathDueCOVID), "n", "y")) 
  
  # Mutate and save desired variables ------------------------------------------------------
  full_num_data_cases <- mortality_og %>% 
    filter(SpCollDt >= ymd(start_date) & SpCollDt <= ymd(end_date)) %>% 
    nrow()

  ids_of_deaths <- mortality_og %>% 
    filter(DeathDueCOVID == "y") %>% 
    pull(unique_num)
    
  mortality_cleaned <- mortality_og %>% 
    mutate(covid_death = ifelse(unique_num %in% ids_of_deaths, "yes", "no"))

  load(file = here("data/positivity-data", "cleaned_process_pcr_data.Rdata"))
  
  
  all_pos <- data.frame(all_pcr_and_zip[["pcr_results_merged"]]) %>% 
    filter(covid_positive == 1)

  
  pos_results_adjusted <- mortality_cleaned %>% 
    select(id = unique_num, death_due_to_covid = DeathDueCOVID, time_days, decades_old, death_date) %>% 
    right_join(y = all_pos, by = "id") %>% 
    select(-c(adj_time_days)) %>% 
    filter(!is.na(death_due_to_covid)) %>% 
    mutate(zip = droplevels(zip))
    
  
  # Read and filter hospital bed data -----------------------------------------------------------
  # OC data
  hos_bed_gov <- read_csv(
    here("data/mortality-data", "covid19hospitalbycounty.csv"),
    na = c("", " "),
    col_types = cols(
      .default = col_skip(),
      county = col_character(),
      todays_date = col_date("%Y-%m-%d"),
      hospitalized_covid_patients = col_double(),
      all_hospital_beds = col_double(),
      icu_covid_confirmed_patients = col_double(),
      icu_suspected_covid_patients = col_double(),
      icu_available_beds = col_double()
    )
  ) %>% 
    filter(county == "Orange") 
  
  
  first_date <- sort(hos_bed_gov$todays_date)[1]
  if(month(params$first_test_date) == month(first_date) & day(params$first_test_date) < day(first_date)) {
    dates_missing <- seq(as.Date(params$first_test_date), first_date, by = "days")
    dates_missing <- dates_missing[-length(dates_missing)]
    num_dm <- length(dates_missing)
    missing_rows <- data.frame(
      rep("Orange", num_dm), 
      dates_missing, 
      rep(NA, num_dm), 
      rep(NA, num_dm), 
      rep(NA, num_dm),
      rep(NA, num_dm),
      rep(NA, num_dm)
    )
    colnames(missing_rows) <- colnames(hos_bed_gov)
    hos_bed_gov <- rbind(missing_rows, hos_bed_gov)
  } else {
    print("Error: fix hospital data dates")
  }

  # Beds were not recorded before 2020-04-20. To fill in missing:
  # For ICU available beds the earliest value is used
  # For percent of beds not used by COVID-19 patients the earliest value was used
  hos_bed_gov <- hos_bed_gov %>% 
    mutate(avail_icu_beds = ifelse(
      is.na(icu_available_beds), 
      params$oc_icu_avail_beds_earliest_val,
      icu_available_beds
    )) %>% 
    mutate(perc_avail_beds = ifelse(
      is.na(hospitalized_covid_patients) | is.na(all_hospital_beds),
      100 * (1 - params$oc_hos_covid_pateients_earliest_val / params$oc_all_hos_bed_earliest_val),
      100 * (1 - hospitalized_covid_patients / all_hospital_beds)
    )) %>% 
    mutate(covid_icu_beds = ifelse(
      is.na(icu_covid_confirmed_patients) | is.na(icu_suspected_covid_patients),
      params$oc_icu_full_beds_earliest_val,
      icu_covid_confirmed_patients + icu_suspected_covid_patients
    )) %>% 
    select(
      test_date = todays_date,
      avail_icu_beds,
      perc_avail_beds,
      covid_icu_beds
    )
  
  hos_bed_gov$adj_perc_avail_beds <- scale(
    hos_bed_gov$perc_avail_beds, 
    center = TRUE, 
    scale = TRUE
  )
  
  hos_bed_gov$adj_avail_icu_beds <- scale(
    hos_bed_gov$avail_icu_beds, 
    center = TRUE, 
    scale = TRUE
  )
  
  hos_bed_gov$adj_covid_icu_beds <- scale(
    hos_bed_gov$covid_icu_beds,
    center = TRUE,
    scale = TRUE
  )
  
  pos_results_merged <- merge(x = pos_results_adjusted, y = hos_bed_gov, by.x = "posted_date", by.y = "test_date")

  
  # Scale time ----------------------------------------------------------------------------------------------
  time_range <- 1:as.integer(difftime(as.Date(end_date), as.Date(start_date), units = "days"))
  adj_time_range <- scale(
    time_range,
    center = TRUE,
    scale = TRUE
  )
  time_data <- data.frame("time_days" = time_range, "adj_time_days" = adj_time_range)
  pos_results_merged <- merge(pos_results_merged, time_data, by = "time_days")

  
  
  # Scale age ----------------------------------------------------------------------------------------------
  pos_results_merged <- pos_results_merged %>% 
    filter(age != 119)
  age_range <- min(pos_results_merged$age):max(pos_results_merged$age)
  adj_age_range <- scale(
    age_range,
    center = TRUE,
    scale = TRUE
  )
  age_data <- data.frame("age" = age_range, "adj_age" = adj_age_range)
  
  pos_results_merged <- merge(pos_results_merged, age_data, by = "age")
  
  
  # Count missing values ------------------------------------------------------------------------------------
  missing_vec <- c(
    full_num_data_cases , # number of observations with test date between start date and end date
    full_num_data_cases - nrow(pos_results_adjusted),  # number of NA for age, sex and zip
    nrow(pos_results_adjusted) - nrow(pos_results_merged) # number of invalid or non Orange County zip codes
  )
  names(missing_vec) <- c(
    "full_num_data_cases", 
    "num_na_cases_removed", 
    "num_bad_zip_cases_removed"
  )
  
  
  list(
    "pos_results_merged" = pos_results_merged, 
    "zip_data_merged" = all_pcr_and_zip$zip_data_merged, 
    "missing" = missing_vec
  )
}

```


```{r load-clean-data}
if (params$reclean_data) {
  pos_pcr_and_zip <- read_pos_pcr(
    file_path = here("data/mortality-data", "all-positive-tests-updated-2020-11-09.csv"),
    start_date = params$first_test_date,
    end_date = params$last_test_date
  )
  save(pos_pcr_and_zip, file = here("data/mortality-data", "cleaned_process_pos_pcr_data.Rdata"))
}
  load(file = here("data/mortality-data", "cleaned_process_pos_pcr_data.Rdata"))
  
  pos_pcr <- data.frame(pos_pcr_and_zip[["pos_results_merged"]])
  pos_zip <- data.frame(pos_pcr_and_zip[["zip_data_merged"]])
  pos_missing <- pos_pcr_and_zip[["missing"]]

```


```{r missing-data-table}
usable_counts <- c(
  pos_missing[1],
  pos_missing[1] - pos_missing[2],
  pos_missing[1] - sum(pos_missing[2:3])
)

tab_obs <- data.frame(
  c(NA, pos_missing[2:3]), 
  usable_counts
)
rownames(tab_obs) <- c(
  "Unmodified",
  "Missing or Inconclusive[note]",
  "Invalid Zip Code"
)

kable(
  tab_obs,
  col.names = c(
    "With Issue", 
    "After Removal"
  ),
  format = "latex",
  caption = paste(
    "Number of positive COVID-19 tests in Orange County from ",
      params$first_test_date, " to ", params$last_test_date,
      ". Includes number removed from data by type of issue for this analysis.",
    sep = "")
  ) %>%
  kable_styling(latex_options = c("HOLD_position")) %>% 
  add_header_above(c(" " = 1, "Number of Tests" = 2), line = TRUE) %>% 
  column_spec(column = 2:3, width = "5.5cm") %>% 
  add_footnote(
    c(
      "Number without conclusive test result or complete demographic information"
      ),
    notation = "symbol",
    threeparttable = TRUE
    )
```


```{r results-functions} 
compute_ci_logistic_glm <- function(model, alpha = 0.05, param_names, include_pvalues = FALSE) {
  #does not include intercept or median income
  model_summary <- summary(model)
  coeffs <- model_summary$coefficients[-1, 1]
  se_robust <- sqrt(diag(sandwich(model)))[-1]
  model_sum <- data.frame("odds" = exp(coeffs),
                          "lower_bound" = exp(coeffs - qnorm(1 - alpha / 2) * se_robust),
                          "upper_bound" = exp(coeffs + qnorm(1 - alpha / 2) * se_robust),
                          "p-value" =  model_summary$coefficients[-1, 4],
                          "param_names" = param_names)
  if(include_pvalues){
    return("model_sum" = model_sum)
  } else {
    return("model_sum" = model_sum[, -4])
  }
}
```


For simplicity let  
$O_{i}$ be the odds of mortality among individuals who tested positive for SARS-CoV-2 in Orange County.  
$\overrightarrow{\beta}_{\text{Race}} = (\beta_{\text{Asian}}, \beta_{\text{Black}}, \beta_{\text{Hispanic}}, \beta_{\text{Native American}}, \beta_{\text{Pacific Islander}}, \beta_{\text{Unknown}})$  
$\overrightarrow{\beta}_{\text{College}} = (\beta_{\text{\% with College Degree Quartile 2}}, \beta_{\text{\% with College Degree Quartile 3}}, \beta_{\text{\% with College Degree Quartile 4}})$  
$\overrightarrow{\beta}_{\text{Insurance}} = (\beta_{\text{\% with Insurance Quartile 2}}, \beta_{\text{\% with Insurance Quartile 3}}, \beta_{\text{\% with Insurance Quartile 4}})$

Model 0:
\begin{equation}
  \label{eq:Model0}
  \begin{split}
    \log(O_i) &= \beta_0 + \beta_{\text{Age}}\text{Age}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}\text{Time}_i\\
		&+ \beta_{\text{\% Hospital Beds Available}}\text{\% Hospital Beds Available}_i,
  \end{split}
\end{equation}
without a random intercept for zip code.

```{r fit-model0}
if (params$refit_model0) {
  fit0 <- glm(
    formula = death_due_to_covid  ~ decades_old + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days + adj_covid_icu_beds,
    family = binomial, 
    data = pos_pcr
    )

  save(fit0, file = here("analysis/mortality-regression-results", "fit0.Rdata"))
} else {
  load(file = here("analysis/mortality-regression-results", "fit0.Rdata"))
}
```


Model 1:
\begin{equation}
  \label{eq:Model1}
  \begin{split}
    \log(O_i) &= \beta_0 + \beta_{\text{Age}}\text{Age}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}\text{Time}_i\\
		&+ \beta_{\text{\% Hospital Beds Available}}\text{\% Hospital Beds Available}_i,
  \end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model1}
if (params$refit_model1) {
  tic()
  fit1 <- glmer(
    formula = death_due_to_covid  ~ decades_old + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days + adj_covid_icu_beds +
              (1 | zip),              
    family = binomial, 
    data = pos_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  #skrrrahh("drummaboy")
  
  save(fit1, file = here("analysis/mortality-regression-results", "fit1.Rdata"))
} else {
  load(file = here("analysis/mortality-regression-results", "fit1.Rdata"))
}

```


```{r bic-table}
compare_bic <- BIC(fit0, fit1) 
row.names(compare_bic) <- c("Model 0", "Model 1")

kable(compare_bic,
      col.names = c("Degrees of Freedom", "BIC"),
      format = "latex",
      caption = "Model comparison using BIC shows do real difference in modeling odds of mortality given tested positive for SARS-CoV-2 in Orange County. Therefore the simpler model, Model 0, was chosen.") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r computes-model-odds-cis}
parameter_names <- c(
  "Age (decades)",
  "Male",
  "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Unknown",
  "2nd Quartile", "3rd Quartile", "4th Quartile",
  " 2nd Quartile", " 3rd Quartile", " 4th Quartile",
  "Population Density (1000ppl/km^2)", 
  "Median Income (std. dev.)",
  "Time (std. dev.)",
  "COVID ICU patients (std. dev.)"
)

summary_fit0 <- compute_ci_logistic_glm(fit0, param_names = parameter_names)
```


```{r fit0-model-results-table}
fm_sum <- summary(fit0)

intercept <- paste0(
  round(exp(fm_sum$coefficients[1, 1]), 3),
  " (",
  round(exp(fm_sum$coefficients[1, 1] - qnorm(0.975) * sqrt(diag(sandwich(fit0)))[1]), 3),
  ",",
  round(exp(fm_sum$coefficients[1, 1] + qnorm(0.975) * sqrt(diag(sandwich(fit0)))[1]), 3),
  ")"
)


fit_final_model <- summary_fit0

rr_fm <- round(fit_final_model[, "odds"], 3)
ci_lb_fm <- round(fit_final_model[, "lower_bound"], 2)
ci_ub_fm <- round(fit_final_model[, "upper_bound"], 2)
#p_values_fm <- fit_final_model$p.value
#p_values_fm <- ifelse(p_values_fm <= 0.001, "0.000", paste(round(p_values_fm, 3)))

reference_group <- c(
  FALSE,
  TRUE, FALSE,
  TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  FALSE,
  FALSE,
  FALSE,
  FALSE
  )

table_fm <- data.frame(matrix(NA, nrow = nrow(fit_final_model) + sum(reference_group), ncol = 3))
colnames(table_fm) <- c("COVID-19 Deaths", "Total", "with (95% CI[note])")
rownames(table_fm) <- c(
  "Age (decades)",
  "Female", "Male",
  "White", "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Unknown",
  "1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile",
  "1st Quartile ", "2nd Quartile ", "3rd Quartile ", "4th Quartile ",
  "Population Density (1000ppl/km^2)",
  "Median Income (std. dev.)",
  "Time (std. dev.)",
  "ICU beds with COVID patients (std. dev.)"
  )

          
non_ref_counter <- 1
for(i in 1:nrow(table_fm)) {
  if (reference_group[i]) {
    #table_fm[i, 1] <- "Reference"
    table_fm[i, 3] <- "Reference"
  } else {
    table_fm[i, 3] <- paste(rr_fm[non_ref_counter], " (", 
                          ci_lb_fm[non_ref_counter], ", ", 
                          ci_ub_fm[non_ref_counter], ")", 
                          sep = "")
    #table_fm[i, 2] <- p_values_fm[non_ref_counter]
    non_ref_counter <- non_ref_counter + 1
  }
}      

variable_counts <- data.frame(matrix(NA,  nrow = nrow(fit_final_model) + sum(reference_group), ncol = 2))
dead_pcr <- pos_pcr[pos_pcr$death_due_to_covid == "y", ]
quant_vars <- c("decades_old", "adj_pop_density", "adj_med_income", "adj_time_days", "adj_avail_icu_beds")
order_quant_vars <- c(1, 19:22)
variable_counts[order_quant_vars, c(1, 2)] <- c("", "")

qual_vars <- c("sex", "race", "adj_perc_bach_quar", "adj_perc_insured_quar")
order_qual_vars <- (1:nrow(table_fm))[-order_quant_vars]

qual_vars_ordered <- c(
  "sex.female", "sex.male", 
  "race.white", "race.asian",
  "race.black or african american", "race.hispanic or latino",
  "race.american indian or alaska native", "race.native hawaiian or other pacific islander",
  "race.unknown",
  "adj_perc_bach_quar.Q1", "adj_perc_bach_quar.Q2",
  "adj_perc_bach_quar.Q3", "adj_perc_bach_quar.Q4",
  "adj_perc_insured_quar.Q1", "adj_perc_insured_quar.Q2",
  "adj_perc_insured_quar.Q3", "adj_perc_insured_quar.Q4"
)
 
calc_freq <- function(df, order_vars) {
  qual_count <- unlist(apply(df, MARGIN = 2, FUN = table))

  qual_count <- qual_count[order_vars]
  qual_freq <- round(100 * qual_count / nrow(df), 2)
  paste(qual_count, " (", qual_freq, "%)", sep = "")
}

variable_counts[order_qual_vars, 2] <- calc_freq(pos_pcr[, qual_vars], qual_vars_ordered)
variable_counts[order_qual_vars, 1] <- calc_freq(dead_pcr[, qual_vars], qual_vars_ordered)

table_fm[, 2] <- variable_counts[, 2]
table_fm[, 1] <- variable_counts[, 1]


kable(
  table_fm, 
  format = "latex", 
  caption = paste(
    "Model 0 regression estimation of adjusted odds ratio of mortality given tested positive for SARS-CoV-2 in Orange County.", 
    sep = ""
    )
  ) %>%
  kable_styling(latex_options = c("HOLD_position"), font_size = 11) %>%
  pack_rows("Sex", 2, 3, bold = FALSE) %>%
  pack_rows("Race", 4, 10, bold = FALSE) %>%
  pack_rows("% with College Degree[note]", 11, 14, bold = FALSE) %>%
  pack_rows("% with Insurance", 15, 18, bold = FALSE) %>% 
  add_header_above(c(" " = 1, "Counts" = 2, "Adjusted Odds Ratio[note]" = 1), line = FALSE) %>% 
  add_footnote(
    c(
      paste0(
        "Model intercept represents odds of death for a white female diagnosed with SARS-CoV-2 in the 0 to 4 age group in a zip code in the first quartile of college degree and insured with the average population density and average number of ICU beds filled with COVID patients in Orange County. The odds of this individual testing dying is estimated to be ",
        intercept,
        collapse = ""
      ), 
      "95% confidence interval computed with robust standard errors",
      "The esimated percent of people with a bachelor's degree, and similarly the estimated percent of people with medical insurance, in an individual's zip code"
      ),
    notation = "symbol"
  )  
```


```{r fit0-model-results-plot, fig.align="center", fig.cap="Logistic model 0 results for odds of mortality among individuals who tested positive for SARS-CoV-2 in Orange County."}

fm_results_with_ref <- rbind(
  summary_fit0[1, ],
  c(1, 0.99, 1.01, paste0(c("Sex", rep(" ", 38), "[ref: Female]"), collapse = "")),
  summary_fit0[2, ],
  c(1, 0.99, 1.01, paste0(c("Race/Ethnicity", rep(" ", 24), "[ref: White]"), collapse = "")),
  summary_fit0[3:8, ],
  c(1, 0.99, 1.01, "% with College Degree [ref: 1st Quartile]"),
  summary_fit0[9:11, ],
  c(1, 0.99, 1.01, paste0(c("% with Insurance", rep(" ", 10), "[ref: 1st Quartile]"), collapse = "")),
  summary_fit0[12:nrow(summary_fit0), ]
)

fm_results_with_ref$odds <- as.double(fm_results_with_ref$odds)
fm_results_with_ref$lower_bound <- as.double(fm_results_with_ref$lower_bound)
fm_results_with_ref$upper_bound <- as.double(fm_results_with_ref$upper_bound)

variable_colors <- c(
  rep("black", 1),
  "white",
  rep("black", 1),
  "white",
  rep("black", 6),
  "white",
  rep("black", 3),
  "white",
  rep("black", 3),
  rep("black", 4)
)

# will need to add colors again to make reference groups white
ggplot(fm_results_with_ref, aes(x = param_names, y = odds), size = 5, color = variable_colors) +
  ylim(min(fm_results_with_ref$lower_bound), max(fm_results_with_ref$upper_bound)) +
  geom_point(color = variable_colors) +
  geom_linerange(aes(ymin = lower_bound, ymax = upper_bound), color = variable_colors) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
  ylab("Adjusted Odds Ratio") +
  coord_flip() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x=element_text(size = 12, colour="black"),
    axis.text.y=element_text(colour="black", hjust = 1),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "black"),
    panel.grid.major = element_line(colour = "gray"),
    panel.background = element_blank(),
    plot.title = element_text(hjust=0)
  ) +
  scale_x_discrete(limits = rev(fm_results_with_ref$param_names)) +
  scale_y_continuous(breaks = seq(0, 2.5, by = 0.5)) +
  ggtitle("Logistic regression results for odds ratio of COVID-19 related mortality given SARS-CoV-2 positive \n in Orange County")
```


