---
title: "Factors associated with mortality among individuals who tested positive for SARS-CoV-2 in Orange County"
output: 
  pdf_document:
    extra_dependencies:
      amsmath 
params:
  reclean_data: FALSE
  refit_model0: FALSE
  refit_model1: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  error = FALSE,
  fig.pos = "H"
)
```


```{r loading-libraries}
# data-wrangling-function chunk
library(tidyverse)
library(here)
library(lubridate)
# missing-data-table chunk
library(knitr)
library(kableExtra)
usepackage_latex("threeparttable")
# fit-model1 chunk
library(sandwich)
library(tictoc)
library(lme4)
#library(BRRR)
library(car) # I don't think I need this anymore
library(mgcv)
```


```{r load-clean-data}
if (params$reclean_data) {
  source(here("analysis", "clean-covid-data.R"))
  
} 

load(usable_cases, file = here("data/cleaned-data", "usable_cases.Rdata"))


pos_pcr <- usable_cases
```


```{r results-functions} 
compute_ci_logistic_glm <- function(model, alpha = 0.05, param_names, include_pvalues = FALSE) {
  #does not include intercept or median income
  model_summary <- summary(model)
  coeffs <- model_summary$coefficients[-1, 1]
  se_robust <- sqrt(diag(sandwich(model)))[-1]
  model_sum <- data.frame("odds" = exp(coeffs),
                          "lower_bound" = exp(coeffs - qnorm(1 - alpha / 2) * se_robust),
                          "upper_bound" = exp(coeffs + qnorm(1 - alpha / 2) * se_robust),
                          "p-value" =  model_summary$coefficients[-1, 4],
                          "param_names" = param_names)
  if(include_pvalues){
    return("model_sum" = model_sum)
  } else {
    return("model_sum" = model_sum[, -4])
  }
}
```


For simplicity let  
$O_{i}$ be the odds of mortality among individuals who tested positive for SARS-CoV-2 in Orange County.  
$\overrightarrow{\beta}_{\text{Race}} = (\beta_{\text{Asian}}, \beta_{\text{Black}}, \beta_{\text{Hispanic}}, \beta_{\text{Native American}}, \beta_{\text{Pacific Islander}}, \beta_{\text{Unknown}})$  
$\overrightarrow{\beta}_{\text{College}} = (\beta_{\text{\% with College Degree Quartile 2}}, \beta_{\text{\% with College Degree Quartile 3}}, \beta_{\text{\% with College Degree Quartile 4}})$  
$\overrightarrow{\beta}_{\text{Insurance}} = (\beta_{\text{\% with Insurance Quartile 2}}, \beta_{\text{\% with Insurance Quartile 3}}, \beta_{\text{\% with Insurance Quartile 4}})$

Model 0:
\begin{equation}
  \label{eq:Model0}
  \begin{split}
    \log(O_i) &= \beta_0 + \beta_{\text{Age}}\text{Age}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{House Crowding}}\text{House Crowding}_i 
		&+ \beta_{\text{Median Income}}\text{Median Income}_i + \beta_{\text{Time}}\text{Time}_i\\
		&+ \beta_{\text{\% Hospital Beds Available}}\text{\% Hospital Beds Available}_i,
  \end{split}
\end{equation}
without a random intercept for zip code.

```{r fit-model0}
if (params$refit_model0) {
  fit0 <- glm(
    formula = death_due_to_covid  ~ decades_old + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + house_crowding + adj_med_income +
              adj_time_days + adj_covid_icu_beds,
    family = binomial, 
    data = pos_pcr
    )

  save(fit0, file = here("analysis/mortality-regression-results", "fit0.Rdata"))
} else {
  load(file = here("analysis/mortality-regression-results", "fit0.Rdata"))
}
```


Model 1:
\begin{equation}
  \label{eq:Model1}
  \begin{split}
    \log(O_i) &= \beta_0 + \beta_{\text{Age}}\text{Age}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{House Crowding}}\text{House Crowding}_i
		&+ \beta_{\text{Median Income}}\text{Median Income}_i + \beta_{\text{Time}}\text{Time}_i\\
		&+ \beta_{\text{\% Hospital Beds Available}}\text{\% Hospital Beds Available}_i,
  \end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model1}
if (params$refit_model1) {
  tic()
  fit1 <- glmer(
    formula = death_due_to_covid  ~ decades_old + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + house_crowding + adj_med_income +
              adj_time_days + adj_covid_icu_beds +
              (1 | zip),              
    family = binomial, 
    data = pos_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  #skrrrahh("drummaboy")
  
  save(fit1, file = here("analysis/mortality-regression-results", "fit1.Rdata"))
} else {
  load(file = here("analysis/mortality-regression-results", "fit1.Rdata"))
}

```


```{r bic-table}
compare_bic <- BIC(fit0, fit1) 
row.names(compare_bic) <- c("Model 0", "Model 1")

kable(compare_bic,
      col.names = c("Degrees of Freedom", "BIC"),
      format = "latex",
      caption = "Model comparison using BIC shows do real difference in modeling odds of mortality given tested positive for SARS-CoV-2 in Orange County. Therefore the simpler model, Model 0, was chosen.") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r computes-model-odds-cis}
parameter_names <- c(
  "Age (decades)",
  "Male",
  "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Unknown",
  "2nd Quartile", "3rd Quartile", "4th Quartile",
  " 2nd Quartile", " 3rd Quartile", " 4th Quartile",
  "Population Density (1000ppl/km^2)", 
  "House Crowding",
  "Median Income (std. dev.)",
  "Time (std. dev.)",
  "COVID ICU patients (std. dev.)"
)

summary_fit0 <- compute_ci_logistic_glm(fit0, param_names = parameter_names)
```


```{r fit0-model-results-table}
fm_sum <- summary(fit0)

intercept <- paste0(
  round(exp(fm_sum$coefficients[1, 1]), 3),
  " (",
  round(exp(fm_sum$coefficients[1, 1] - qnorm(0.975) * sqrt(diag(sandwich(fit0)))[1]), 3),
  ",",
  round(exp(fm_sum$coefficients[1, 1] + qnorm(0.975) * sqrt(diag(sandwich(fit0)))[1]), 3),
  ")"
)


fit_final_model <- summary_fit0

rr_fm <- round(fit_final_model[, "odds"], 3)
ci_lb_fm <- round(fit_final_model[, "lower_bound"], 2)
ci_ub_fm <- round(fit_final_model[, "upper_bound"], 2)
#p_values_fm <- fit_final_model$p.value
#p_values_fm <- ifelse(p_values_fm <= 0.001, "0.000", paste(round(p_values_fm, 3)))

reference_group <- c(
  FALSE,
  TRUE, FALSE,
  TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  FALSE,
  FALSE,
  FALSE,
  FALSE,
  FALSE
  )

table_fm <- data.frame(matrix(NA, nrow = nrow(fit_final_model) + sum(reference_group), ncol = 3))
colnames(table_fm) <- c("COVID-19 Deaths", "Total", "with (95% CI[note])")
rownames(table_fm) <- c(
  "Age (decades)",
  "Female", "Male",
  "White", "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Unknown",
  "1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile",
  "1st Quartile ", "2nd Quartile ", "3rd Quartile ", "4th Quartile ",
  "Population Density (1000ppl/km^2)",
  "House Crowding",
  "Median Income (std. dev.)",
  "Time (std. dev.)",
  "ICU beds with COVID patients (std. dev.)"
  )

          
non_ref_counter <- 1
for(i in 1:nrow(table_fm)) {
  if (reference_group[i]) {
    #table_fm[i, 1] <- "Reference"
    table_fm[i, 3] <- "Reference"
  } else {
    table_fm[i, 3] <- paste(rr_fm[non_ref_counter], " (", 
                          ci_lb_fm[non_ref_counter], ", ", 
                          ci_ub_fm[non_ref_counter], ")", 
                          sep = "")
    #table_fm[i, 2] <- p_values_fm[non_ref_counter]
    non_ref_counter <- non_ref_counter + 1
  }
}      

variable_counts <- data.frame(matrix(NA,  nrow = nrow(fit_final_model) + sum(reference_group), ncol = 2))
dead_pcr <- pos_pcr[pos_pcr$death_due_to_covid == "y", ]
quant_vars <- c("decades_old", "adj_pop_density", "house_crowding", "adj_med_income", "adj_time_days", "adj_avail_icu_beds")
order_quant_vars <- c(1, 19:23)
variable_counts[order_quant_vars, c(1, 2)] <- c("", "")

qual_vars <- c("sex", "race", "adj_perc_bach_quar", "adj_perc_insured_quar")
order_qual_vars <- (1:nrow(table_fm))[-order_quant_vars]

qual_vars_ordered <- c(
  "sex.female", "sex.male", 
  "race.white", "race.asian",
  "race.black or african american", "race.hispanic or latino",
  "race.american indian or alaska native", "race.native hawaiian or other pacific islander",
  "race.unknown",
  "adj_perc_bach_quar.Q1", "adj_perc_bach_quar.Q2",
  "adj_perc_bach_quar.Q3", "adj_perc_bach_quar.Q4",
  "adj_perc_insured_quar.Q1", "adj_perc_insured_quar.Q2",
  "adj_perc_insured_quar.Q3", "adj_perc_insured_quar.Q4"
)
 
calc_freq <- function(df, order_vars) {
  qual_count <- unlist(apply(df, MARGIN = 2, FUN = table))

  qual_count <- qual_count[order_vars]
  qual_freq <- round(100 * qual_count / nrow(df), 2)
  paste(qual_count, " (", qual_freq, "%)", sep = "")
}

variable_counts[order_qual_vars, 2] <- calc_freq(pos_pcr[, qual_vars], qual_vars_ordered)
variable_counts[order_qual_vars, 1] <- calc_freq(dead_pcr[, qual_vars], qual_vars_ordered)

table_fm[, 2] <- variable_counts[, 2]
table_fm[, 1] <- variable_counts[, 1]


kable(
  table_fm, 
  format = "latex", 
  caption = paste(
    "Model 0 regression estimation of adjusted odds ratio of mortality given tested positive for SARS-CoV-2 in Orange County.", 
    sep = ""
    )
  ) %>%
  kable_styling(latex_options = c("HOLD_position"), font_size = 11) %>%
  pack_rows("Sex", 2, 3, bold = FALSE) %>%
  pack_rows("Race", 4, 10, bold = FALSE) %>%
  pack_rows("% with College Degree[note]", 11, 14, bold = FALSE) %>%
  pack_rows("% with Insurance", 15, 18, bold = FALSE) %>% 
  add_header_above(c(" " = 1, "Counts" = 2, "Adjusted Odds Ratio[note]" = 1), line = FALSE) %>% 
  add_footnote(
    c(
      paste0(
        "Model intercept represents odds of death for a white female diagnosed with SARS-CoV-2 in the 0 to 4 age group in a zip code in the first quartile of college degree and insured with the average population density and average number of ICU beds filled with COVID patients in Orange County. The odds of this individual testing dying is estimated to be ",
        intercept,
        collapse = ""
      ), 
      "95% confidence interval computed with robust standard errors",
      "The esimated percent of people with a bachelor's degree, and similarly the estimated percent of people with medical insurance, in an individual's zip code"
      ),
    notation = "symbol"
  )  
```


```{r fit0-model-results-plot, fig.align="center", fig.cap="Logistic model 0 results for odds of mortality among individuals who tested positive for SARS-CoV-2 in Orange County."}

fm_results_with_ref <- rbind(
  summary_fit0[1, ],
  c(1, 0.99, 1.01, paste0(c("Sex", rep(" ", 38), "[ref: Female]"), collapse = "")),
  summary_fit0[2, ],
  c(1, 0.99, 1.01, paste0(c("Race/Ethnicity", rep(" ", 24), "[ref: White]"), collapse = "")),
  summary_fit0[3:8, ],
  c(1, 0.99, 1.01, "% with College Degree [ref: 1st Quartile]"),
  summary_fit0[9:11, ],
  c(1, 0.99, 1.01, paste0(c("% with Insurance", rep(" ", 10), "[ref: 1st Quartile]"), collapse = "")),
  summary_fit0[12:nrow(summary_fit0), ]
)

fm_results_with_ref$odds <- as.double(fm_results_with_ref$odds)
fm_results_with_ref$lower_bound <- as.double(fm_results_with_ref$lower_bound)
fm_results_with_ref$upper_bound <- as.double(fm_results_with_ref$upper_bound)

variable_colors <- c(
  rep("black", 1),
  "white",
  rep("black", 1),
  "white",
  rep("black", 6),
  "white",
  rep("black", 3),
  "white",
  rep("black", 3),
  rep("black", 5)
)

# will need to add colors again to make reference groups white
ggplot(fm_results_with_ref, aes(x = param_names, y = odds), size = 5, color = variable_colors) +
  ylim(min(fm_results_with_ref$lower_bound), max(fm_results_with_ref$upper_bound)) +
  geom_point(color = variable_colors) +
  geom_linerange(aes(ymin = lower_bound, ymax = upper_bound), color = variable_colors) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
  ylab("Adjusted Odds Ratio") +
  coord_flip() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x=element_text(size = 12, colour="black"),
    axis.text.y=element_text(colour="black", hjust = 1),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "black"),
    panel.grid.major = element_line(colour = "gray"),
    panel.background = element_blank(),
    plot.title = element_text(hjust=0)
  ) +
  scale_x_discrete(limits = rev(fm_results_with_ref$param_names)) +
  scale_y_continuous(breaks = seq(0, 2.5, by = 0.5)) +
  ggtitle("Logistic regression results for odds ratio of COVID-19 related mortality given SARS-CoV-2 positive \n in Orange County")
```


