---
title: "Factors associated with seroprevelance of SARS-CoV-2 in Orange County"
output: 
  pdf_document:
    extra_dependencies:
      amsmath 
params:
  reclean_data: FALSE
  refit_model0: FALSE
  refit_model1: FALSE
  refit_model2: FALSE
  refit_model3: FALSE
  refit_model4: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE, 
                      error = FALSE,
                      fig.pos = "H")
```


```{r loading-libraries}
# data-wrangling-function chunk
library(tidyverse)
library(here)
library(lubridate)
# missing-data-table chunk
library(knitr)
library(kableExtra)
usepackage_latex("threeparttable")
# fit-model1 chunk
library(sandwich)
library(tictoc)
library(lme4)
#(BRRR)
library(car) # I don't think I need this anymore
library(mgcv)
```


```{r load-clean-data}
if (params$reclean_data) {
  source(here("analysis", "clean-covid-data.R"))
  
} 

load(sero_results_merged, file = here("data/cleaned-data", "sero_results_merged.Rdata"))

all_sero <- sero_results_merged
```


```{r results-functions} 
compute_ci_logistic_glm <- function(model, alpha = 0.05, param_names, include_pvalues = FALSE) {
  #does not include intercept or median income
  model_summary <- summary(model)
  coeffs <- model_summary$coefficients[-1, 1]
  se_robust <- sqrt(diag(sandwich(model)))[-1]
  model_sum <- data.frame("odds" = exp(coeffs),
                          "lower_bound" = exp(coeffs - qnorm(1 - alpha / 2) * se_robust),
                          "upper_bound" = exp(coeffs + qnorm(1 - alpha / 2) * se_robust),
                          "p-value" =  model_summary$coefficients[-1, 4],
                          "param_names" = param_names)
  if(include_pvalues){
    return("model_sum" = model_sum)
  } else {
    return("model_sum" = model_sum[, -4])
  }
}
```


For simplicity let  
$O_i$ be the odds of testing positive for COVID-19 in Orange County.  
$\overrightarrow{\beta}_{\text{Age Group}} = (\beta_{\text{Age[18-24]}}, \beta_{\text{Age[25-29]}}, \beta_{\text{Age[30-34]}}, \beta_{\text{Age[35-39]}}, \beta_{\text{Age[40-44]}}, \beta_{\text{Age[45-49]}}, \beta_{\text{Age[50-54]}}, \\ \beta_{\text{Age[55-59]}}, \beta_{\text{Age[60-64]}}, \beta_{\text{Age[65-69]}}, \beta_{\text{Age[70-74]}}, \beta_{\text{Age[75-79]}}, \beta_{\text{Age[80-84]}}, \beta_{\text{Age[85+]}})$  
$\overrightarrow{\beta}_{\text{Race}} = (\beta_{\text{Asian}}, \beta_{\text{Black}}, \beta_{\text{Hispanic}}, \beta_{\text{PacificIslander}}, \beta_{\text{Unknown}})$  
$\overrightarrow{\beta}_{\text{College}} = (\beta_{\text{\% with College Degree Quartile 2}}, \beta_{\text{\% with College Degree Quartile 3}}, \beta_{\text{\% with College Degree Quartile 4}})$  
$\overrightarrow{\beta}_{\text{Insurance}} = (\beta_{\text{\% with Insurance Quartile 2}}, \beta_{\text{\% with Insurance Quartile 3}}, \beta_{\text{\% with Insurance Quartile 4}})$


Model 0:
\begin{equation}
\label{eq:Model0}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{House Crowding}}\text{House Crowding}_i
		+ \beta_{\text{Median Income}}\text{Median Income}_i,
	\end{split}
\end{equation}
without a random intercept for zip code.


```{r fit-model0}
if (params$refit_model0) {
  fit0 <- glm(
    formula = covid_pos ~ age_grp + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + house_crowding + adj_med_income,
    family = binomial, 
    data = all_sero
    )

  save(fit0, file = here("analysis/testing-seropositive-regression-results", "fit0.Rdata"))
} else {
  load(file = here("analysis/testing-seropositive-regression-results", "fit0.Rdata"))
}
```


Model 1:
\begin{equation}
\label{eq:Model1}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{House Crowding}}\text{House Crowding}_i + \beta_{\text{Median Income}}\text{Median Income}_i,
	\end{split}
\end{equation}
with a random intercept for zip code.


```{r fit-model1}
if (params$refit_model1) {
  tic()
  fit1 <- glmer(
    formula = covid_pos ~ age_grp + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + house_crowding + adj_med_income +
              (1 | zip),              
    family = binomial, 
    data = all_sero,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()

  # tic()
  # ss <- getME(fit1a, c("theta","fixef"))
  # fit1b <- update(
  #   fit1a,
  #   start = ss,
  #   control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e6))
  # )
  # toc()
  # fit1 <- fit1b
  
  save(fit1, file = here("analysis/testing-seropositive-regression-results", "fit1.Rdata"))
} else {
  load(file = here("analysis/testing-seropositive-regression-results", "fit1.Rdata"))
}
```


Model 2:
\begin{equation}
\label{eq:Model2}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{House Crowding}}\text{House Crowding}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\
		&+ \beta_{\text{\% of Zip Code SARS-CoV-2+}}\text{\% of Zip Code SARS-CoV-2+}_i,
	\end{split}
\end{equation}
without a random intercept for zip code.


```{r fit-seropos-modeled-with-num-cases}
if (params$refit_model2) {
  fit2 <- glm(
    formula = covid_pos ~ age_grp + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + house_crowding + adj_med_income +
              I(perc_zip_covid_pos/10),
    family = binomial, 
    data = all_sero
    )

  save(fit2, file = here("analysis/testing-seropositive-regression-results", "fit2.Rdata"))
} else {
  load(file = here("analysis/testing-seropositive-regression-results", "fit2.Rdata"))
}
```


Model 3:
\begin{equation}
\label{eq:Model3}
	\begin{split}
		\log(O_i) &= \beta_0 + \beta_{\text{\% of Zip Code SARS-CoV-2+}}\text{\% of Zip Code SARS-CoV-2+}_i,
	\end{split}
\end{equation}
without a random intercept for zip code.


```{r fit-seropos-modeled-by-num-cases}
if (params$refit_model3) {
  fit3 <- glm(
    formula = covid_pos ~ I(perc_zip_covid_pos/10),
    family = binomial, 
    data = all_sero
    )

  save(fit3, file = here("analysis/testing-seropositive-regression-results", "fit3.Rdata"))
} else {
  load(file = here("analysis/testing-seropositive-regression-results", "fit3.Rdata"))
}
```


Model 4:
\begin{equation}
\label{eq:Model4}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \beta_{\text{\% of Zip Code SARS-CoV-2+}}\text{\% of Zip Code SARS-CoV-2+}_i,
	\end{split}
\end{equation}
without a random intercept for zip code.


```{r fit-seropos-modeled-without-zip-vars}
if (params$refit_model4) {
  fit4 <- glm(
    formula = covid_pos ~ age_grp + sex + race + 
              I(perc_zip_covid_pos/10),
    family = binomial, 
    data = all_sero
    )

  save(fit4, file = here("analysis/testing-seropositive-regression-results", "fit4.Rdata"))
} else {
  load(file = here("analysis/testing-seropositive-regression-results", "fit4.Rdata"))
}
```


```{r compute-model-odds-cis}
parameter_names <- c(
  "25-29", "30-34", "35-39", "40-49", "50-59", "60-69", "70-79", "80+", 
  "Male",
  "Asian", "Black", "Hispanic", "Pacific Islander", "Unknown",
  "2nd Quartile", "3rd Quartile", "4th Quartile",
  " 2nd Quartile", " 3rd Quartile", " 4th Quartile",
  "Population Density (1000ppl/km^2)", 
  "House Crowding",
  "Median Income (std. dev.)"
)

parameter_names4 <- c(
  "25-29", "30-34", "35-39", "40-49", "50-59", "60-69", "70-79", "80+", 
  "Male",
  "Asian", "Black", "Hispanic", "Pacific Islander", "Unknown",
  "% of Zip SARS-CoV-2+ (10%)"
)

summary_fit0 <- compute_ci_logistic_glm(fit0, param_names = parameter_names)
summary_fit2 <- compute_ci_logistic_glm(fit2, param_names = c(parameter_names, "% of Zip SARS-CoV-2+ (10%)"))
summary_fit3 <- compute_ci_logistic_glm(fit3, param_names = "% of Zip SARS-CoV-2+ (10%)")
summary_fit4 <- compute_ci_logistic_glm(fit4, param_names = parameter_names4)
```


```{r bic-table}
compare_bic <- BIC(fit0, fit1, fit2, fit3, fit4) 
row.names(compare_bic) <- c("Model 0", "Model 1", "Model 2", "Model 3", "Model 4")

kable(compare_bic,
      col.names = c("Degrees of Freedom", "BIC"),
      format = "latex",
      caption = "Model comparison using BIC shows negligible difference in modeling odds of testing sero-positive for COVID-19 in Orange County. Therefore the simpler model, Model 0, was chosen.") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r fit0-model-results-table}
fm_sum <- summary(fit0)

intercept <- paste0(
  round(exp(fm_sum$coefficients[1, 1]), 3),
  " (",
  round(exp(fm_sum$coefficients[1, 1] - qnorm(0.975) * sqrt(diag(sandwich(fit0)))[1]), 3),
  ",",
  round(exp(fm_sum$coefficients[1, 1] + qnorm(0.975) * sqrt(diag(sandwich(fit0)))[1]), 3),
  ")"
)


fit_final_model <- summary_fit0

rr_fm <- round(fit_final_model[, "odds"], 3)
ci_lb_fm <- round(fit_final_model[, "lower_bound"], 2)
ci_ub_fm <- round(fit_final_model[, "upper_bound"], 2)
#p_values_fm <- fit_final_model$p.value
#p_values_fm <- ifelse(p_values_fm <= 0.001, "0.000", paste(round(p_values_fm, 3)))

reference_group <- c(
  TRUE, rep(FALSE, 8),
  TRUE, FALSE,
  TRUE, rep(FALSE, 5),
  TRUE, rep(FALSE, 3),
  TRUE, rep(FALSE, 3),
  FALSE,
  FALSE,
  FALSE
)

table_fm <- data.frame(matrix(NA, nrow = nrow(fit_final_model) + sum(reference_group), ncol = 3))
colnames(table_fm) <- c("SARS-CoV-2+", "Total", "with (95% CI[note])")
rownames(table_fm) <- c(
  "18-24", "25-29", "30-34", "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Female", "Male",
  "White", "Asian", "Black", "Hispanic", "Pacific Islander", "Unknown",
  "1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile",
  "1st Quartile ", "2nd Quartile ", "3rd Quartile ", "4th Quartile ",
  "Population Density (1000ppl/km^2)",
  "House Crowding",
  "Median Income (std. dev.)"
)

          
non_ref_counter <- 1
for(i in 1:nrow(table_fm)) {
  if (reference_group[i]) {
    #table_fm[i, 1] <- "Reference"
    table_fm[i, 3] <- "Reference"
  } else {
    table_fm[i, 3] <- paste(rr_fm[non_ref_counter], " (", 
                          ci_lb_fm[non_ref_counter], ", ", 
                          ci_ub_fm[non_ref_counter], ")", 
                          sep = "")
    #table_fm[i, 2] <- p_values_fm[non_ref_counter]
    non_ref_counter <- non_ref_counter + 1
  }
}      

variable_counts <- data.frame(matrix(NA,  nrow = nrow(fit_final_model) + sum(reference_group), ncol = 2))
sero_pos <- all_sero[all_sero$covid_pos == "yes", ]
quant_vars <- c("adj_pop_density", "house_crowding", "adj_med_income")
order_quant_vars <- (nrow(table_fm) - 2):nrow(table_fm)
variable_counts[order_quant_vars, c(1, 2)] <- rep("", length(quant_vars))

qual_vars <- c("age_grp", "sex", "race", "adj_perc_bach_quar", "adj_perc_insured_quar")
order_qual_vars <- (1:nrow(table_fm))[-order_quant_vars]

qual_vars_ordered <- c(
  "age_grp.18-24", "age_grp.25-29", "age_grp.30-34", "age_grp.35-39", 
  "age_grp.40-49", "age_grp.50-59","age_grp.60-69", "age_grp.70-79", "age_grp.80+",
  "sex.female", "sex.male", 
  "race.white", "race.asian", "race.black", "race.hispanic", 
  "race.islander", "race.unknown",
  "adj_perc_bach_quar.Q1", "adj_perc_bach_quar.Q2",
  "adj_perc_bach_quar.Q3", "adj_perc_bach_quar.Q4",
  "adj_perc_insured_quar.Q1", "adj_perc_insured_quar.Q2",
  "adj_perc_insured_quar.Q3", "adj_perc_insured_quar.Q4"
)
 
calc_freq <- function(df, order_vars) {
  qual_count <- unlist(apply(df, MARGIN = 2, FUN = table))

  qual_count <- qual_count[order_vars]
  qual_freq <- round(100 * qual_count / nrow(df), 2)
  paste(qual_count, " (", qual_freq, "%)", sep = "")
}

variable_counts[order_qual_vars, 2] <- calc_freq(all_sero[, qual_vars], qual_vars_ordered)
variable_counts[order_qual_vars, 1] <- calc_freq(sero_pos[, qual_vars], qual_vars_ordered)

table_fm[, 2] <- variable_counts[, 2]
table_fm[, 1] <- variable_counts[, 1]

kable(
  table_fm, 
  format = "latex", 
  caption = paste(
    "Model 0 regression estimation of adjusted odds ratio of testing sero-positive for SARS-CoV-2 in Orange County.", 
    sep = ""
    )
  ) %>%
  kable_styling(latex_options = c("HOLD_position"), font_size = 11) %>%
  pack_rows("Age", 1, 9, bold = FALSE) %>% 
  pack_rows("Sex", 10, 11, bold = FALSE) %>%
  pack_rows("Race[note]", 12, 17, bold = FALSE) %>%
  pack_rows("% with College Degree[note]", 18, 21, bold = FALSE) %>%
  pack_rows("% with Insurance", 22, 24, bold = FALSE) %>% 
  add_header_above(c(" " = 1, "Counts" = 2, "Adjusted Odds Ratio[note]" = 1), line = FALSE) %>% 
  add_footnote(
    c(
      paste0(
        "Model intercept represents odds of testing sero-positive for SARS-CoV-2 for a white female diagnosed with SARS-CoV-2 in the 18-24 age group in a zip code in the first quartile of college degree and insured with the average population density in Orange County. The odds of this individual testing sero-positive is estimated to be ",
        intercept,
        collapse = ""
      ), 
      "95% confidence interval computed with robust standard errors",
      "Native American/Native Alaskan race group not included in analysis due to lack of data, no individual of this race group tested seropositive.",
      "The esimated percent of people with a bachelor's degree, and similarly the estimated percent of people with medical insurance, in an individual's zip code"
    ),
    notation = "symbol"
  )  
```


```{r fit0-model-results-plot, fig.align="center", fig.cap="Logistic model 0 results for odds of testing sero-positive for SARS-CoV-2 in Orange County. Pasific Islander adjusted odds ratio omitted due to width of confidence interval (1.05, 15.14)."}

fm_results_with_ref <- rbind(
  c(1, 0.99, 1.01, paste0(c("Age", rep(" ", 39), "[ref: 18-24]"), collapse = "")),
  summary_fit0[1:8, ],
  c(1, 0.99, 1.01, paste0(c("Sex", rep(" ", 38), "[ref: Female]"), collapse = "")),
  summary_fit0[9, ],
  c(1, 0.99, 1.01, paste0(c("Race", rep(" ", 38), "[ref: White]"), collapse = "")),
  summary_fit0[10:14, ],
  c(1, 0.99, 1.01, "% with College Degree [ref: 1st Quartile]"),
  summary_fit0[15:17, ],
  c(1, 0.99, 1.01, paste0(c("% with Insurance", rep(" ", 10), "[ref: 1st Quartile]"), collapse = "")),
  summary_fit0[18:nrow(summary_fit0), ]
)

fm_results_with_ref$odds <- as.double(fm_results_with_ref$odds)
fm_results_with_ref$lower_bound <- as.double(fm_results_with_ref$lower_bound)
fm_results_with_ref$upper_bound <- as.double(fm_results_with_ref$upper_bound)

variable_colors <- c(
  "white",
  rep("black", 8),
  "white",
  rep("black", 1),
  "white",
  rep("black", 4),
  "white",
  rep("black", 3),
  "white",
  rep("black", 3),
  rep("black", 3)
)

fm_results_with_ref <- fm_results_with_ref %>% 
  filter(param_names != "Pacific Islander")

ggplot(fm_results_with_ref, aes(x = param_names, y = odds), size = 5, color = variable_colors) +
  ylim(min(fm_results_with_ref$lower_bound), max(fm_results_with_ref$upper_bound)) +
  geom_point(color = variable_colors) +
  geom_linerange(aes(ymin = lower_bound, ymax = upper_bound), color = variable_colors) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
  ylab("Adjusted Odds Ratio") +
  coord_flip() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x=element_text(size = 12, colour="black"),
    axis.text.y=element_text(colour="black", hjust = 1),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "black"),
    panel.grid.major = element_line(colour = "gray"),
    panel.background = element_blank(),
    plot.title = element_text(hjust=0)
  ) +
  scale_x_discrete(limits = rev(fm_results_with_ref$param_names)) +
  scale_y_continuous(breaks = seq(0, max(fm_results_with_ref$upper_bound), by = 0.5)) +
  ggtitle("Results of SARS-CoV-2 seropositivity logisitic regression")
```


```{r fit2-model-results-table}
fm_sum <- summary(fit2)

intercept <- paste0(
  round(exp(fm_sum$coefficients[1, 1]), 3),
  " (",
  round(exp(fm_sum$coefficients[1, 1] - qnorm(0.975) * sqrt(diag(sandwich(fit2)))[1]), 3),
  ",",
  round(exp(fm_sum$coefficients[1, 1] + qnorm(0.975) * sqrt(diag(sandwich(fit2)))[1]), 3),
  ")"
)


fit_final_model <- summary_fit2

rr_fm <- round(fit_final_model[, "odds"], 3)
ci_lb_fm <- round(fit_final_model[, "lower_bound"], 2)
ci_ub_fm <- round(fit_final_model[, "upper_bound"], 2)
#p_values_fm <- fit_final_model$p.value
#p_values_fm <- ifelse(p_values_fm <= 0.001, "0.000", paste(round(p_values_fm, 3)))

reference_group <- c(
  TRUE, rep(FALSE, 8),
  TRUE, FALSE,
  TRUE, rep(FALSE, 5),
  TRUE, rep(FALSE, 3),
  TRUE, rep(FALSE, 3),
  FALSE,
  FALSE,
  FALSE,
  FALSE
)

table_fm <- data.frame(matrix(NA, nrow = nrow(fit_final_model) + sum(reference_group), ncol = 3))
colnames(table_fm) <- c("SARS-CoV-2+", "Total", "with (95% CI[note])")
rownames(table_fm) <- c(
  "18-24", "25-29", "30-34", "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Female", "Male",
  "White", "Asian", "Black", "Hispanic", "Pacific Islander", "Unknown",
  "1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile",
  "1st Quartile ", "2nd Quartile ", "3rd Quartile ", "4th Quartile ",
  "Population Density (1000ppl/km^2)",
  "House Crowding",
  "Median Income (std. dev.)",
  "% of Zip Code SARS-CoV-2+[note]"
)

          
non_ref_counter <- 1
for(i in 1:nrow(table_fm)) {
  if (reference_group[i]) {
    #table_fm[i, 1] <- "Reference"
    table_fm[i, 3] <- "Reference"
  } else {
    table_fm[i, 3] <- paste(rr_fm[non_ref_counter], " (", 
                          ci_lb_fm[non_ref_counter], ", ", 
                          ci_ub_fm[non_ref_counter], ")", 
                          sep = "")
    #table_fm[i, 2] <- p_values_fm[non_ref_counter]
    non_ref_counter <- non_ref_counter + 1
  }
}      

variable_counts <- data.frame(matrix(NA,  nrow = nrow(fit_final_model) + sum(reference_group), ncol = 2))
sero_pos <- all_sero[all_sero$covid_pos == "yes", ]
quant_vars <- c("adj_pop_density", "house_crowding", "adj_med_income", "perc_zip_covid_pos")
order_quant_vars <- (nrow(table_fm) - 3):nrow(table_fm)
variable_counts[order_quant_vars, c(1, 2)] <- rep("", length(quant_vars))

qual_vars <- c("age_grp", "sex", "race", "adj_perc_bach_quar", "adj_perc_insured_quar")
order_qual_vars <- (1:nrow(table_fm))[-order_quant_vars]

qual_vars_ordered <- c(
  "age_grp.18-24", "age_grp.25-29", "age_grp.30-34", "age_grp.35-39", 
  "age_grp.40-49", "age_grp.50-59","age_grp.60-69", "age_grp.70-79", "age_grp.80+",
  "sex.female", "sex.male", 
  "race.white", "race.asian", "race.black", "race.hispanic", 
  "race.islander", "race.unknown",
  "adj_perc_bach_quar.Q1", "adj_perc_bach_quar.Q2",
  "adj_perc_bach_quar.Q3", "adj_perc_bach_quar.Q4",
  "adj_perc_insured_quar.Q1", "adj_perc_insured_quar.Q2",
  "adj_perc_insured_quar.Q3", "adj_perc_insured_quar.Q4"
)
 
calc_freq <- function(df, order_vars) {
  qual_count <- unlist(apply(df, MARGIN = 2, FUN = table))

  qual_count <- qual_count[order_vars]
  qual_freq <- round(100 * qual_count / nrow(df), 2)
  paste(qual_count, " (", qual_freq, "%)", sep = "")
}

variable_counts[order_qual_vars, 2] <- calc_freq(all_sero[, qual_vars], qual_vars_ordered)
variable_counts[order_qual_vars, 1] <- calc_freq(sero_pos[, qual_vars], qual_vars_ordered)

table_fm[, 2] <- variable_counts[, 2]
table_fm[, 1] <- variable_counts[, 1]

kable(
  table_fm, 
  format = "latex", 
  caption = paste(
    "Model 2 regression estimation of adjusted odds ratio of testing sero-positive for SARS-CoV-2 in Orange County.", 
    sep = ""
    )
  ) %>%
  kable_styling(latex_options = c("HOLD_position"), font_size = 11) %>%
  pack_rows("Age", 1, 9, bold = FALSE) %>% 
  pack_rows("Sex", 10, 11, bold = FALSE) %>%
  pack_rows("Race[note]", 12, 17, bold = FALSE) %>%
  pack_rows("% with College Degree[note]", 18, 21, bold = FALSE) %>%
  pack_rows("% with Insurance", 22, 24, bold = FALSE) %>% 
  add_header_above(c(" " = 1, "Counts" = 2, "Adjusted Odds Ratio[note]" = 1), line = FALSE) %>% 
  add_footnote(
    c(
      paste0(
        "Model intercept represents odds of testing sero-positive for SARS-CoV-2 for a white female diagnosed with SARS-CoV-2 in the 18-24 age group in a zip code in the first quartile of college degree and insured with the average population density, and average percent of SARS-CoV-2 positive individuals in Orange County. The odds of this individual testing sero-positive is estimated to be ",
        intercept,
        collapse = ""
      ), 
      "95% confidence interval computed with robust standard errors",
      "Native American/Native Alaskan race group not included in analysis due to lack of data, no individual of this race group tested seropositive",
      "The esimated percent of people with a bachelor's degree, and similarly the estimated percent of people with medical insurance, in an individual's zip code",
      "Number of individuals who tested positive in individual's zip code reported to OC Public Health Department from March 1st to August 16th, divided by estimated population of zipcode"
    ),
    notation = "symbol"
  )  
```


```{r fit2-model-results-plot, fig.align="center", fig.cap="Logistic model 2 results for odds of testing sero-positive for SARS-CoV-2 in Orange County. Pasific Islander adjusted odds ratio omitted due to width of confidence interval (1.05, 15.14)."}

fm_results_with_ref <- rbind(
  c(1, 0.99, 1.01, paste0(c("Age", rep(" ", 39), "[ref: 18-24]"), collapse = "")),
  summary_fit2[1:8, ],
  c(1, 0.99, 1.01, paste0(c("Sex", rep(" ", 38), "[ref: Female]"), collapse = "")),
  summary_fit2[9, ],
  c(1, 0.99, 1.01, paste0(c("Race", rep(" ", 38), "[ref: White]"), collapse = "")),
  summary_fit2[10:14, ],
  c(1, 0.99, 1.01, "% with College Degree [ref: 1st Quartile]"),
  summary_fit2[15:17, ],
  c(1, 0.99, 1.01, paste0(c("% with Insurance", rep(" ", 10), "[ref: 1st Quartile]"), collapse = "")),
  summary_fit2[18:nrow(summary_fit2), ]
)

fm_results_with_ref$odds <- as.double(fm_results_with_ref$odds)
fm_results_with_ref$lower_bound <- as.double(fm_results_with_ref$lower_bound)
fm_results_with_ref$upper_bound <- as.double(fm_results_with_ref$upper_bound)

variable_colors <- c(
  "white",
  rep("black", 8),
  "white",
  rep("black", 1),
  "white",
  rep("black", 4),
  "white",
  rep("black", 3),
  "white",
  rep("black", 3),
  rep("black", 4)
)

fm_results_with_ref <- fm_results_with_ref %>% 
  filter(param_names != "Pacific Islander")

ggplot(fm_results_with_ref, aes(x = param_names, y = odds), size = 5, color = variable_colors) +
  ylim(min(fm_results_with_ref$lower_bound), max(fm_results_with_ref$upper_bound)) +
  geom_point(color = variable_colors) +
  geom_linerange(aes(ymin = lower_bound, ymax = upper_bound), color = variable_colors) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
  ylab("Adjusted Odds Ratio") +
  coord_flip() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x=element_text(size = 12, colour="black"),
    axis.text.y=element_text(colour="black", hjust = 1),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "black"),
    panel.grid.major = element_line(colour = "gray"),
    panel.background = element_blank(),
    plot.title = element_text(hjust=0)
  ) +
  scale_x_discrete(limits = rev(fm_results_with_ref$param_names)) +
  scale_y_continuous(breaks = seq(0, max(fm_results_with_ref$upper_bound), by = 0.5)) +
  ggtitle("Logistic regression results for odds ratio of testing sero-positive for SARS-CoV-2 in Orange County")
```


```{r fit3-model-results-plot, fig.align="center", fig.cap="Logistic model 3 results for odds of testing sero-positive for SARS-CoV-2 in Orange County modeled only by percent of their zip code that has tested positive for SARS-CoV-2"}

fm_results_with_ref <- rbind(
  summary_fit3[1, ]
)

fm_results_with_ref$odds <- as.double(fm_results_with_ref$odds)
fm_results_with_ref$lower_bound <- as.double(fm_results_with_ref$lower_bound)
fm_results_with_ref$upper_bound <- as.double(fm_results_with_ref$upper_bound)

variable_colors <- c(
  "black"
)


ggplot(fm_results_with_ref, aes(x = param_names, y = odds), size = 5, color = variable_colors) +
  ylim(min(fm_results_with_ref$lower_bound), max(fm_results_with_ref$upper_bound)) +
  geom_point(color = variable_colors) +
  geom_linerange(aes(ymin = lower_bound, ymax = upper_bound), color = variable_colors) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
  ylab("Adjusted Odds Ratio") +
  coord_flip() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x=element_text(size = 12, colour="black"),
    axis.text.y=element_text(colour="black", hjust = 1),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "black"),
    panel.grid.major = element_line(colour = "gray"),
    panel.background = element_blank(),
    plot.title = element_text(hjust=0)
  ) +
  scale_x_discrete(limits = rev(fm_results_with_ref$param_names)) +
  scale_y_continuous(breaks = seq(0, max(fm_results_with_ref$upper_bound), by = 0.25)) +
  ggtitle("Results of SARS-CoV-2 seropositivity logisitic regression")
```


```{r fit4-model-results-table}
fm_sum <- summary(fit4)

intercept <- paste0(
  round(exp(fm_sum$coefficients[1, 1]), 3),
  " (",
  round(exp(fm_sum$coefficients[1, 1] - qnorm(0.975) * sqrt(diag(sandwich(fit4)))[1]), 3),
  ",",
  round(exp(fm_sum$coefficients[1, 1] + qnorm(0.975) * sqrt(diag(sandwich(fit4)))[1]), 3),
  ")"
)


fit_final_model <- summary_fit4

rr_fm <- round(fit_final_model[, "odds"], 3)
ci_lb_fm <- round(fit_final_model[, "lower_bound"], 2)
ci_ub_fm <- round(fit_final_model[, "upper_bound"], 2)
#p_values_fm <- fit_final_model$p.value
#p_values_fm <- ifelse(p_values_fm <= 0.001, "0.000", paste(round(p_values_fm, 3)))

reference_group <- c(
  TRUE, rep(FALSE, 8),
  TRUE, FALSE,
  TRUE, rep(FALSE, 5),
  FALSE
)

table_fm <- data.frame(matrix(NA, nrow = nrow(fit_final_model) + sum(reference_group), ncol = 3))
colnames(table_fm) <- c("SARS-CoV-2+", "Total", "with (95% CI[note])")
rownames(table_fm) <- c(
  "18-24", "25-29", "30-34", "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Female", "Male",
  "White", "Asian", "Black", "Hispanic", "Pacific Islander", "Unknown",
  "% of Zip Code SARS-CoV-2+[note]"
)

          
non_ref_counter <- 1
for(i in 1:nrow(table_fm)) {
  if (reference_group[i]) {
    #table_fm[i, 1] <- "Reference"
    table_fm[i, 3] <- "Reference"
  } else {
    table_fm[i, 3] <- paste(rr_fm[non_ref_counter], " (", 
                          ci_lb_fm[non_ref_counter], ", ", 
                          ci_ub_fm[non_ref_counter], ")", 
                          sep = "")
    #table_fm[i, 2] <- p_values_fm[non_ref_counter]
    non_ref_counter <- non_ref_counter + 1
  }
}      

variable_counts <- data.frame(matrix(NA,  nrow = nrow(fit_final_model) + sum(reference_group), ncol = 2))
sero_pos <- all_sero[all_sero$covid_pos == "yes", ]
quant_vars <- c("adj_pop_density", "adj_med_income", "perc_zip_covid_pos")
order_quant_vars <- nrow(table_fm)
variable_counts[order_quant_vars, c(1, 2)] <- c("", "")

qual_vars <- c("age_grp", "sex", "race", "adj_perc_bach_quar", "adj_perc_insured_quar")
order_qual_vars <- (1:nrow(table_fm))[-order_quant_vars]

qual_vars_ordered <- c(
  "age_grp.18-24", "age_grp.25-29", "age_grp.30-34", "age_grp.35-39", 
  "age_grp.40-49", "age_grp.50-59","age_grp.60-69", "age_grp.70-79", "age_grp.80+",
  "sex.female", "sex.male", 
  "race.white", "race.asian", "race.black", "race.hispanic", 
  "race.islander", "race.unknown"
)
 
calc_freq <- function(df, order_vars) {
  qual_count <- unlist(apply(df, MARGIN = 2, FUN = table))

  qual_count <- qual_count[order_vars]
  qual_freq <- round(100 * qual_count / nrow(df), 2)
  paste(qual_count, " (", qual_freq, "%)", sep = "")
}

variable_counts[order_qual_vars, 2] <- calc_freq(all_sero[, qual_vars], qual_vars_ordered)
variable_counts[order_qual_vars, 1] <- calc_freq(sero_pos[, qual_vars], qual_vars_ordered)

table_fm[, 2] <- variable_counts[, 2]
table_fm[, 1] <- variable_counts[, 1]

kable(
  table_fm, 
  format = "latex", 
  caption = paste(
    "Model 4 regression estimation of adjusted odds ratio of testing sero-positive for SARS-CoV-2 in Orange County.", 
    sep = ""
    )
  ) %>%
  kable_styling(latex_options = c("HOLD_position"), font_size = 11) %>%
  pack_rows("Age", 1, 9, bold = FALSE) %>% 
  pack_rows("Sex", 10, 11, bold = FALSE) %>%
  pack_rows("Race[note]", 12, 17, bold = FALSE) %>%
  add_header_above(c(" " = 1, "Counts" = 2, "Adjusted Odds Ratio[note]" = 1), line = FALSE) %>% 
  add_footnote(
    c(
      paste0(
        "Model intercept represents odds of testing sero-positive for SARS-CoV-2 for a white female diagnosed with SARS-CoV-2 in the 18-24 age group in a zip code in the first quartile of college degree and insured with the average population density, and average percent of SARS-CoV-2 positive individuals in Orange County. The odds of this individual testing sero-positive is estimated to be ",
        intercept,
        collapse = ""
      ), 
      "95% confidence interval computed with robust standard errors",
      "Native American/Native Alaskan race group not included in analysis due to lack of data, no individual of this race group tested seropositive",
      "Number of individuals who tested positive in individual's zip code reported to OC Public Health Department from March 1st to August 16th, divided by estimated population of zipcode"
    ),
    notation = "symbol"
  )  
```


```{r fit4-model-results-plot, fig.align="center", fig.cap="Logistic model 4 results for odds of testing sero-positive for SARS-CoV-2 in Orange County. Pasific Islander adjusted odds ratio omitted due to width of confidence interval (1.05, 15.14)."}

fm_results_with_ref <- rbind(
  c(1, 0.99, 1.01, paste0(c("Age", rep(" ", 39), "[ref: 18-24]"), collapse = "")),
  summary_fit4[1:8, ],
  c(1, 0.99, 1.01, paste0(c("Sex", rep(" ", 38), "[ref: Female]"), collapse = "")),
  summary_fit4[9, ],
  c(1, 0.99, 1.01, paste0(c("Race", rep(" ", 38), "[ref: White]"), collapse = "")),
  summary_fit4[10:nrow(summary_fit4), ]
)

fm_results_with_ref$odds <- as.double(fm_results_with_ref$odds)
fm_results_with_ref$lower_bound <- as.double(fm_results_with_ref$lower_bound)
fm_results_with_ref$upper_bound <- as.double(fm_results_with_ref$upper_bound)

variable_colors <- c(
  "white",
  rep("black", 8),
  "white",
  rep("black", 1),
  "white",
  rep("black", 4),
  rep("black", 1)
)

fm_results_with_ref <- fm_results_with_ref %>% 
  filter(param_names != "Pacific Islander")

ggplot(fm_results_with_ref, aes(x = param_names, y = odds), size = 5, color = variable_colors) +
  ylim(min(fm_results_with_ref$lower_bound), max(fm_results_with_ref$upper_bound)) +
  geom_point(color = variable_colors) +
  geom_linerange(aes(ymin = lower_bound, ymax = upper_bound), color = variable_colors) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
  ylab("Adjusted Odds Ratio") +
  coord_flip() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x=element_text(size = 12, colour="black"),
    axis.text.y=element_text(colour="black", hjust = 1),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "black"),
    panel.grid.major = element_line(colour = "gray"),
    panel.background = element_blank(),
    plot.title = element_text(hjust=0)
  ) +
  scale_x_discrete(limits = rev(fm_results_with_ref$param_names)) +
  scale_y_continuous(breaks = seq(0, max(fm_results_with_ref$upper_bound), by = 0.5)) +
  ggtitle("Results of SARS-CoV-2 seropositivity logisitic regression (with incidence)")
```

