---
title: "Factors associated with testing positive for SARS-CoV-2 in Orange County"
output: 
  pdf_document:
    extra_dependencies:
      amsmath 
params:
  reclean_data: FALSE
  refit_model1: FALSE
  refit_model2: FALSE
  refit_model3: FALSE
  refit_model4: FALSE
  refit_model5: FALSE
  refit_model6: FALSE
  refit_model7: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      error = FALSE,
                      fig.pos = "H")
```


```{r loading-libraries}
# data-wrangling-function chunk
library(tidyverse)
library(here)
library(lubridate)
# missing-data-table chunk
library(knitr)
library(kableExtra)
# fit-model1 chunk
library(tictoc)
library(lme4)
library(car) # I don't think I need this anymore
library(mgcv)
```


```{r load-clean-data}
if (params$reclean_data) {
  source(here("analysis", "clean-covid-data.R"))
  
} 

load(file = here("data/cleaned-data", "usable_tests.Rdata"))

all_pcr <- usable_tests
```



```{r compute-ci-gam-logistic-function}
compute_ci_gam_logistic <- function(model, alpha = 0.05, param_names) {
  #does not include intercept or median income
  model_summary <- summary(model)
  in_med_inc <- length(model_summary$p.coeff)
  coeffs <- model_summary$p.coeff[-c(1, in_med_inc)]
  se <- sqrt(diag(vcov(model, unconditional = TRUE)))[2:(length(coeffs) + 1)]
  model_sum <- data.frame(
    "odds" = exp(coeffs),
    "lower_bound" = exp(coeffs - qnorm(1 - alpha / 2) * se),
    "upper_bound" = exp(coeffs + qnorm(1 - alpha / 2) * se),
    "p-value" =  model_summary$p.pv[-c(1, in_med_inc)],
    "param_names" = param_names
  )
  
  return("model_sum" = model_sum)
}
```

For simplicity let  
$O_i$ be the odds of testing positive for SARS-CoV-2 in Orange County.  
$\overrightarrow{\beta}_{\text{Age Group}} = (\beta_{\text{Age[5-9]}}, \beta_{\text{Age[10-14]}}, \beta_{\text{Age[15-19]}}, \beta_{\text{Age[20-24]}}, \beta_{\text{Age[25-29]}}, \beta_{\text{Age[30-34]}}, \beta_{\text{Age[35-39]}}\\ \beta_{\text{Age[40-49]}},  \beta_{\text{Age[50-59]}}, \beta_{\text{Age[60-69]}}, \beta_{\text{Age[70-79]}}, \beta_{\text{Age[80+]}})$  
$\overrightarrow{\beta}_{\text{Race}} = (\beta_{\text{Asian}}, \beta_{\text{Black}}, \beta_{\text{Hispanic}}, \beta_{\text{NativeAmerican}}, \beta_{\text{PacificIslander}}, \beta_{\text{OtherRace}}, \beta_{\text{Unknown}})$  
$\overrightarrow{\beta}_{\text{College}} = (\beta_{\text{\% with College Degree Quartile 2}}, \beta_{\text{\% with College Degree Quartile 3}}, \beta_{\text{\% with College Degree Quartile 4}})$  
$\overrightarrow{\beta}_{\text{Insurance}} = (\beta_{\text{\% with Insurance Quartile 2}}, \beta_{\text{\% with Insurance Quartile 3}}, \beta_{\text{\% with Insurance Quartile 4}})$


Model 1:
\begin{equation}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Gender}}\text{Gender}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}\text{Time}_i,
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model1}
if (params$refit_model1) {
  tic()
  fit_time_lin <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days +
              (1 | zip),              
    family = binomial, 
    data = all_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  # 
  # 2730.94 sec elapsed
  # updated and saved 2021-06-04
  save(fit_time_lin, file = here("analysis/testing-positive-regression-results", "fit_time_lin.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_lin.Rdata"))
}
```

Model 2:
\begin{equation}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Gender}}\text{Gender}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}\text{Time}_i + \beta_{\text{Time2}}\text{Time}^2_i,
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model2}
if (params$refit_model2) {
  tic()
  fit_time_quad <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              I(adj_time_days) + I(adj_time_days^2) + 
              (1 | zip),              
    family = binomial, 
    data = all_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  # 
  # 2916.06 sec elapsed
  # Updated 2020-06-04
  save(fit_time_quad, file = here("analysis/testing-positive-regression-results", "fit_time_quad.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_quad.Rdata"))
}
```

Model 3:
\begin{equation}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Gender}}\text{Gender}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}\text{Time}_i + \beta_{\text{Time2}}\text{Time}^2_i, \\
		&+ \beta_{\text{Interaction1}}\text{Median Income}_i\times\text{Time}_i + \beta_{\text{Interaction2}}\text{Median Income}_i\times\text{Time}^2_i,
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model3}
if (params$refit_model3) {
  tic()
  fit_time_quad_inter <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days + I(adj_time_days^2) + 
              adj_time_days:adj_med_income + I(adj_time_days^2):adj_med_income +
              (1 | zip),              
    family = binomial, 
    data = all_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  # 2903.28 sec elapsed
  # updated 2021-06-04
  save(fit_time_quad_inter, file = here("analysis/testing-positive-regression-results", "fit_time_quad_inter.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_quad_inter.Rdata"))
}
```

Model 4:  
Generalized additive model
\begin{equation}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Gender}}\text{Gender}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}f(\text{Time}_i),
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model4}
if (params$refit_model4) {
  tic()
  fit_time_gam <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density  + adj_med_income +
              s(time_days, bs = "ts", k = -1) + 
              s(zip, bs = "re"),              
    family = binomial, 
    data = all_pcr,
    method = "REML",
    gamma = 1.5
    )
  toc()
  # 
  # updated and saved 2021-06-04
  # 915.72 sec elapsed
  save(fit_time_gam, file = here("analysis/testing-positive-regression-results", "fit_time_gam.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_gam.Rdata"))
}
```

Model 5:  
Generalized additive model
\begin{equation}
		\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Gender}}\text{Gender}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}f(\text{Time}_i) + \beta_{\text{Interaction}}\text{Median Income}_i\times f(\text{Time}_i),
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model5, results="hide"}
if (params$refit_model5) {
  tic()
  fit_time_gam_inter <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density  + adj_med_income +
              s(time_days) +
              ti(time_days, adj_med_income, bs = "ts") + 
              s(zip, bs = "re"),              
    family = binomial, 
    data = all_pcr,
    method = "REML",
    gamma = 1.5
    )
  toc()
  # 
# 1889.62 sec elapsed
# updated and saved 2021-06-04
  save(fit_time_gam_inter, file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter.Rdata"))
}
 
parameter_names <- c(
  "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", 
  "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Male",
  "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Unknown",
  "2nd Quartile", "3rd Quartile", "4th Quartile",
  " 2nd Quartile", " 3rd Quartile", " 4th Quartile",
  "Population Density"
  )

fit_time_gam_inter_summary1 <- compute_ci_gam_logistic(
  model = fit_time_gam_inter,
  param_names = parameter_names
)

gam_smooth_var_ci <- gam.vcomp(fit_time_gam_inter)
```


Model 6:  
Generalized additive model
\begin{equation}
		\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Gender}}\text{Gender}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}f(\text{Time}_i) + \beta_{\text{Interaction}}\text{Median Income}_i\times f(\text{Time}_i),
	\end{split}
\end{equation}
without a random intercept for zip code.

```{r fit-model6, results="hide"}
if (params$refit_model6) {
  tic()
  fit6 <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density  + adj_med_income +
              s(time_days) +
              ti(time_days, adj_med_income, bs = "ts"),
    family = binomial, 
    data = all_pcr,
    method = "REML",
    gamma = 1.5
    )
  toc()
  # 
# 139.25 sec elapsed
# updated and saved 2021-06-04
  save(fit6, file = here("analysis/testing-positive-regression-results", "fit6.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit6.Rdata"))
}
```


Model 7:  
Generalized additive model
\begin{equation}
		\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Gender}}\text{Gender}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{House Crowding}}\text{House Crowding}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}f(\text{Time}_i) + \beta_{\text{Interaction}}\text{Median Income}_i\times f(\text{Time}_i),
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model7, results="hide"}
if (params$refit_model7) {
  tic()
  fit_time_gam_inter_house <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + house_crowding + adj_med_income +
              s(time_days) +
              ti(time_days, adj_med_income, bs = "ts") + 
              s(zip, bs = "re"),              
    family = binomial, 
    data = all_pcr,
    method = "REML",
    gamma = 1.5
    )
  toc()
  # 
# 1862.25 sec elapsed
# updated and saved 2021-06-04
  save(fit_time_gam_inter_house, file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter_house.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter_house.Rdata"))
}
 
parameter_names <- c(
  "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", 
  "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Male",
  "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Unknown",
  "2nd Quartile", "3rd Quartile", "4th Quartile",
  " 2nd Quartile", " 3rd Quartile", " 4th Quartile",
  "Population Density",
  "House Crowding"
  )

fit_time_gam_inter_house_summary1 <- compute_ci_gam_logistic(
  model = fit_time_gam_inter_house,
  param_names = parameter_names
)

gam_smooth_var_ci <- gam.vcomp(fit_time_gam_inter_house)
```


```{r bic-table}
compare_bic <- BIC(fit_time_lin, fit_time_quad, fit_time_quad_inter, fit_time_gam, fit_time_gam_inter, fit6, fit_time_gam_inter_house) 
row.names(compare_bic) <- c("Model 1", "Model 2", "Model 3", "Model 4", "Model 5", "Model 6", "Model 7")

kable(compare_bic,
      col.names = c("Degrees of Freedom", "BIC"),
      format = "latex",
      caption = "Model comparison using BIC supports Model 5 to model odds of testing positive for COVID-19 in Orange County.") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r final-model-results-table}
fit_final_model <- fit_time_gam_inter_house_summary1

rr_fm <- round(fit_final_model[, "odds"], 3)
ci_lb_fm <- round(fit_final_model[, "lower_bound"], 2)
ci_ub_fm <- round(fit_final_model[, "upper_bound"], 2)
#p_values_fm <- fit_final_model$p.value
#p_values_fm <- ifelse(p_values_fm <= 0.001, "0.000", paste(round(p_values_fm, 3)))
# The 5 accounts for the reference groups
table_fm <- data.frame(matrix(NA, nrow = nrow(fit_final_model) + 5, ncol = 3))
colnames(table_fm) <- c("SARS-CoV-2+", "Total", "with (95% CI[note])")
rownames(table_fm) <- c(
  "0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", 
  "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Female", "Male",
  "White", "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Unknown",
  "1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile",
  "1st Quartile ", "2nd Quartile ", "3rd Quartile ", "4th Quartile ",
  "Population Density (1000ppl/km^2)",
  "House Crowding Index"
  )
reference_group <- c(
  TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,
  TRUE, FALSE,
  TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  FALSE,
  FALSE
  )
          
non_ref_counter <- 1
for(i in 1:nrow(table_fm)) {
  if (reference_group[i]) {
    #table_fm[i, 1] <- "Reference"
    table_fm[i, 3] <- "Reference"
  } else {
    table_fm[i, 3] <- paste(rr_fm[non_ref_counter], " (", 
                          ci_lb_fm[non_ref_counter], ", ", 
                          ci_ub_fm[non_ref_counter], ")", 
                          sep = "")
    #table_fm[i, 2] <- p_values_fm[non_ref_counter]
    non_ref_counter <- non_ref_counter + 1
  }
}      

variable_counts <- data.frame(matrix(NA,  nrow = nrow(fit_final_model) + 5, ncol = 2))
all_pcr_ppl <- all_pcr[!duplicated(all_pcr$id), ]
cov_pos_pcr_ppl <- all_pcr_ppl[all_pcr_ppl$covid_positive == 1, ]
quant_vars <- c("pop_density", "house_crowding")
order_quant_vars <- c(nrow(table_fm) - 1, nrow(table_fm))
variable_counts[order_quant_vars, c(1, 2)] <- c("", "")

qual_vars <- c("age_group", "sex", "race", "adj_perc_bach_quar", "adj_perc_insured_quar")
order_qual_vars <- (1:nrow(table_fm))[-order_quant_vars]

qual_vars_ordered <- c(
  "age_group.0-4","age_group.5-9", "age_group.10-14", "age_group.15-19",
  "age_group.20-24", "age_group.25-29", "age_group.30-34", "age_group.35-39",
  "age_group.40-49", "age_group.50-59", "age_group.60-69", "age_group.70-79",
  "age_group.80+", 
  "sex.female", "sex.male", 
  "race.white", "race.asian",
  "race.black or african american", "race.hispanic or latino",
  "race.american indian or alaska native", "race.native hawaiian or other pacific islander",
  "race.unknown",
  "adj_perc_bach_quar.Q1", "adj_perc_bach_quar.Q2",
  "adj_perc_bach_quar.Q3", "adj_perc_bach_quar.Q4",
  "adj_perc_insured_quar.Q1", "adj_perc_insured_quar.Q2",
  "adj_perc_insured_quar.Q3", "adj_perc_insured_quar.Q4"
  )
 
calc_freq <- function(df, order_vars) {
  qual_count <- unlist(apply(df, MARGIN = 2, FUN = table))

  qual_count <- qual_count[order_vars]
  qual_freq <- round(100 * qual_count / nrow(df), 2)
  paste(qual_count, " (", qual_freq, "%)", sep = "")
}

variable_counts[order_qual_vars, 2] <- calc_freq(all_pcr_ppl[, qual_vars], qual_vars_ordered)
variable_counts[order_qual_vars, 1] <- calc_freq(cov_pos_pcr_ppl[, qual_vars], qual_vars_ordered)

table_fm[, 2] <- variable_counts[, 2]
table_fm[, 1] <- variable_counts[, 1]

fm_sum <- summary(fit_time_gam)
intercept <- paste0(
  round(exp(fm_sum$p.coeff[1]), 3),
  " (",
  round(exp(fm_sum$p.coeff[1] - qnorm(0.975) * fm_sum$se[1]), 2),
  ",",
  round(exp(fm_sum$p.coeff[1] + qnorm(0.975) * fm_sum$se[1]), 2),
  ")"
)


kable(
  table_fm, 
  format = "latex", 
  caption = paste(
    "Model 5 logistic regression results for linear fixed effects of odds of testing positive for SARS-CoV-2 in Orange County. This table excludes the coefficient for median income due to the interaction between median income and time.", 
    sep = ""
    )
  ) %>%
  kable_styling(latex_options = c("HOLD_position"), font_size = 11) %>%
  pack_rows("Age", 1, 13, bold = FALSE) %>%
  pack_rows("Gender", 14,15, bold = FALSE) %>%
  pack_rows("Race/ethnicity", 16, 23, bold = FALSE) %>%
  pack_rows("% with College Degree[note]", 24, 27, bold = FALSE) %>%
  pack_rows("% with Insurance", 28, 31, bold = FALSE) %>% 
  add_header_above(c(" " = 1, "Counts" = 2, "Adjusted Odds Ratio[note]" = 1), line = FALSE) %>% 
  add_footnote(
    c(
      paste0(
        "Adjusted for all covariates listed plus zip code estimated median income and time of test in days. Model intercept represents odds of a white female in the 0 to 4 age group in a zip code in the first quartile of college degree and insured with the average population density and no house crowding. The odds of this individual testing positive for COVID-19 is estimated to be ",
        intercept
      ), 
      "95% Confidence Interval",
      "Esimated: percent of people with a bachelor's degree, percent of people with medical insurance, population density, and house crowding index in an individual's zip code "
      ),
    notation = "symbol"
  )  
```


```{r final-model-results-coeff-plot, fig.align="center", fig.cap="Model 5 logistic regression results for linear fixed effects of odds of testing positive for SARS-CoV-2 in Orange County. This plot excludes the coefficient for median income due to the interaction between median income and time. Percent with college degree is the estimated associated change in odds of testing positive for SARS-CoV-2 for an individual living in a zip code where the estimate percent of the population with a bachelor's degree is in a given quartile relative to an individual from a zip code in the first. Percent insured representing estimated percent of zip code population with medical insurance is similarly defined."}

fm_results_with_ref <- rbind(
  c(1, 0.99, 1.01, 0, paste0(c("Age", rep(" ", 34), "[ref: 0-4 years]"), collapse = "")),
  fit_time_gam_inter_house_summary1[1:12, ],
  c(1, 0.99, 1.01, 0, paste0(c("Gender", rep(" ", 33), "[ref: Female]"), collapse = "")),
  fit_time_gam_inter_house_summary1[13, ],
  c(1, 0.99, 1.01, 0, paste0(c("Race/ethnicity", rep(" ", 25), "[ref: White]"), collapse = "")),
  fit_time_gam_inter_house_summary1[14:19, ],
  c(1, 0.99, 1.01, 0, "% with College Degree [ref: 1st Quartile]"),
  fit_time_gam_inter_house_summary1[20:22, ],
  c(1, 0.99, 1.01, 0, paste0(c("% with Insurance", rep(" ", 10), "[ref: 1st Quartile]"), collapse = "")),
  fit_time_gam_inter_house_summary1[23:25, ],
  c(as.character(fit_time_gam_inter_house_summary1[26, -5]), "Population Density (1000ppl/km^2)"),
  c(as.character(fit_time_gam_inter_house_summary1[27, -5]), "House Crowding")
)

fm_results_with_ref$odds <- as.double(fm_results_with_ref$odds)
fm_results_with_ref$lower_bound <- as.double(fm_results_with_ref$lower_bound)
fm_results_with_ref$upper_bound <- as.double(fm_results_with_ref$upper_bound)

variable_colors <- c(
  "white",
  rep("black", 12),
  "white",
  rep("black", 1),
  "white",
  rep("black", 6),
  "white",
  rep("black", 3),
  "white",
  rep("black", 5)
)


ggplot(fm_results_with_ref, aes(x = param_names, y = odds), size = 5) +
  ylim(min(fm_results_with_ref$lower_bound), max(fm_results_with_ref$upper_bound)) +
  geom_point(color = variable_colors) +
  geom_linerange(aes(ymin = lower_bound, ymax = upper_bound), color = variable_colors) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
  ylab("Adjusted Odds Ratio") +
  coord_flip() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x=element_text(size = 12, colour="black"),
    axis.text.y=element_text(colour="black", hjust = 1),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "black"),
    panel.grid.major = element_line(colour = "gray"),
    panel.background = element_blank(),
    plot.title = element_text(hjust=0)
  ) +
  scale_x_discrete(limits = rev(fm_results_with_ref$param_names)) +
  scale_y_continuous(breaks = seq(0, 2.5, by = 0.5)) +
  ggtitle("Generalized additive logistic regression results for odds ratio of testing SARS-CoV-2 positive \n in Orange County")
```


```{r final-model-results-interaction-plot, fig.align="center", fig.width=7, fig.height=7, fig.cap="Three dimentional visualization of interaction between time and median income from logistic model 7. SARS-CoV-2 test results in Orange County from March 1st to August 16th. Large difference visible in early March between individuals in zip codes with relatively high median incomes and those in zip codes with relatively small median incomes; this difference diminishes as time increase. Gap is due to lack of zip codes with median income in interval, the extreme median income is Newport Coast with zip code 92657."}
plot(fit_time_gam_inter_house, scheme = 1, select = 2, 
     xlab = "Time (std. dev.)",
     ylab = "Median Income (std. dev.)",
     main = "Interaction") 
```




