---
title: "Factors associated with testing positive for SARS-CoV-2 in Orange County"
output: 
  pdf_document:
    extra_dependencies:
      amsmath 
params:
  reclean_data: FALSE
  first_test_date: "2020-03-01"
  last_test_date: "2020-08-16"
  refit_model1: FALSE
  refit_model2: FALSE
  refit_model3: FALSE
  refit_model4: FALSE
  refit_model5: FALSE
  refit_model6: FALSE
  refit_model7: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      error = FALSE,
                      fig.pos = "H")
```


```{r loading-libraries}
# data-wrangling-function chunk
library(tidyverse)
library(here)
library(lubridate)
library(data.table)
# missing-data-table chunk
library(knitr)
library(kableExtra)
# fit-model1 chunk
library(tictoc)
library(lme4)
library(BRRR)
library(car) # I don't think I need this anymore
library(mgcv)
```


```{r data-wrangling-function}
# Test results need to be categorized ------------------------------------------------------
negative_test_synonyms <- c(
  "not detected",
  "negative",
  "coronavirus 2019 novel not detected",
  "negative (qualifier value)",
  "not detected (qualifier value)",
  "sars cov-2 negative",
  "undetected",
  "inst_negative",
  "neg-see report",
  "sars-cov-2 rna not detected by naa",
  "none detected", 
  "not detected in pooled specimen"
  )

positive_test_synonyms <- c(
  "detected",
  "coronavirus 2019 novel positive",
  "positive",
  "positive (qualifier value)",
  "sars cov-2 positive",
  "detected (qualifier value)",
  "presumptive pos",
  "positive for 2019-ncov",
  "presumptive positive",
  "coronavirus 2019 novel presumptive pos",
  "coronavirus 2019 novel detected",
  "yes",
  "coronavirus 2019 novel",
  "presumptive positive for 2019-ncov",
  "sars cov-2 presumptive pos",
  "presumptive pos. for 2019-ncov",
  "presumptive positive (qualifier value)",
  "presumptive detected",
  "reactive",
  "sars-cov-2",
  "not detected in pooled specimen (qualifier value)"
  )

other_test_synonyms <- c(
  "inconclusive",
  "indeterminate",
  "specimen unsatisfactory",
  "invalid",
  "test not performed",
  "not provided (qualifier value)",
  "see comment",
  "tnp",
  "coronavirus 2019 novel inconclusive",
  "not tested",
  "phoned results (and readback confirmed) to:",
  "see note",
  "clotted",
  "coronavirus 2019 novel unsatisfactory",
  "equivocal",
  "non reactive",
  "result comments",
  "sars cov-2 inconclusive",
  "test not done",
  "test not perf",
  "not pregnant",
  "biofiresarsneg",
  "equivocal result",
  "coronavirus 2019 novel inconcluside",
  "unsatisfactory",
  "undefined",
  "*corrected report* by",
  "specimen unsatifactory for evaluation",
  "warning....please disregard results.",
  "presumptive result to be confirmed",
  "indeterminate (qualifier value)", 
  "invalid result", 
  "specimen unsatisfactory for evaluation",
  "acinetobacter baumannii (organism)",
  "carbapenem resistant pseudomonas aeruginosa",
  "enterobacter cloacae complex (organism)",
  "enterococcus faecalis",
  "genus enterococcus",
  "interpretive information: 2019 novel coronavirus sars-cov-2 by pcr",
  "multiple drug-resistant serratia marcescens",
  "specimen received mislabeled"
  )


read_all_pcr <- function(file_path, start_date, end_date) {
  pcr_results_original <- read_csv(
    file_path,
    col_types = cols(
      .default = col_skip(),
      PersonId = col_character(),
      Age = col_integer(),
      Sex = col_character(),
      Ethnicity = col_character(),
      Race = col_character(),
      Specimen.Collected.Date = col_date("%m-%d-%Y"),
      Resulted.Organism = col_character(),
      Zip = col_character()
      )
    ) 
  
  full_num_data_cases <- pcr_results_original %>% 
    filter(Specimen.Collected.Date >= ymd(start_date) & Specimen.Collected.Date <= ymd(end_date)) %>% 
    nrow()
  
  full_num_id_cases <- pcr_results_original %>% 
    filter(Specimen.Collected.Date >= ymd(start_date) & Specimen.Collected.Date <= ymd(end_date)) %>% 
    select(PersonId) %>% 
    unique() %>% 
    nrow()

  pcr_results_original$Race[is.na(pcr_results_original$Race)] <- "Unknown"
  pcr_results_original$Ethnicity[is.na(pcr_results_original$Ethnicity)] <- "Unknown"

  hispanic_race_unknown <- (
    (pcr_results_original$Race == "Other" & 
     pcr_results_original$Ethnicity == "Hispanic or Latino") |
    (pcr_results_original$Race == "Unknown" & 
     pcr_results_original$Ethnicity == "Hispanic or Latino") |
    (pcr_results_original$Race == "Multiple Races" & 
     pcr_results_original$Ethnicity == "Hispanic or Latino") 
    )
  
  non_hispanic_unknown <- (
    (pcr_results_original$Race == "Unknown" & 
     pcr_results_original$Ethnicity != "Hispanic or Latino") |
    (pcr_results_original$Race == "Multiple Races" & 
     pcr_results_original$Ethnicity != "Hispanic or Latino")
    )
  
  pcr_results_original_new_race <- data.frame(
    pcr_results_original, 
    "race1" = str_to_lower(pcr_results_original$Race)
    )
  pcr_results_original_new_race$race1[hispanic_race_unknown] <- "hispanic or latino"
  pcr_results_original_new_race$race1[non_hispanic_unknown] <- "unknown"
  
  pcr_results_adjusted <- pcr_results_original_new_race %>%
    filter(!is.na(Resulted.Organism)) %>%
    mutate(test_result = fct_collapse(
      str_to_lower(Resulted.Organism),
      negative = negative_test_synonyms,
      positive = positive_test_synonyms,
      unknown = other_test_synonyms) 
      ) %>%
    mutate(sex = fct_collapse(
      str_to_lower(Sex),
      male = "m",
      female = "f",
      unknown = c("d", "g", "i", "tf", "tm", "u")
      )) %>%
    mutate(race = factor(
      race1, 
      levels = c(
        "white",
        "asian",
        "black or african american",
        "hispanic or latino",
        "american indian or alaska native",
        "native hawaiian or other pacific islander",
        "other",
        "unknown"
        ),
      )) %>% 
    mutate(
      time_days = as.integer(round(difftime(
        Specimen.Collected.Date, 
        start_date, 
        units = "days"
        )))
      ) %>%
    select(
      id = PersonId, 
      posted_date = Specimen.Collected.Date, 
      time_days,
      test_result, 
      age = Age,
      sex,
      race,
      ethnicity = Ethnicity,
      zip = Zip,
      ) %>%
    filter((posted_date >= ymd(start_date)) & (posted_date <= ymd(end_date))) %>%
    filter(test_result != "unknown") %>%
    filter(sex != "unknown") %>%
    mutate(zip = str_sub(zip, end = 5)) %>%
    drop_na() %>% 
    group_by(id) %>%
    arrange(posted_date) %>%
    ungroup()
  
  
  if(length(levels(pcr_results_adjusted$test_result)) != 3) warning("New test result category not accounted for.")

  pcr_results_adjusted$covid_positive <- ifelse( pcr_results_adjusted$test_result == "positive", 1, 0)
  
  
  # Extract df of all observations with an id that has inconsistencies for the demographic variables (age, sex, race)
  pcr_rep_id <- pcr_results_adjusted[(
    duplicated(pcr_results_adjusted$id, fromLast=TRUE) | 
    duplicated(pcr_results_adjusted$id)
    ), ]

  pcr_rep_id2 <- pcr_rep_id %>%
    group_by(id) %>% 
    mutate(reasonable_ages = diff(range(age)) <= 1) %>%
    mutate(identical_race = n_distinct(race) == 1) %>%
    mutate(identical_sex = n_distinct(sex) == 1) %>%
    ungroup()
  
  pcr_inconsistent <- data.frame(
    pcr_rep_id2[(
      !pcr_rep_id2$reasonable_ages | 
      !pcr_rep_id2$identical_sex | 
      !pcr_rep_id2$identical_race
      ), ]
    )
  
  inconsistent_counts <- c(
    length(unique(pcr_inconsistent$id[(!pcr_inconsistent$reasonable_ages)])),
    length(unique(pcr_inconsistent$id[(!pcr_inconsistent$identical_sex)])),
    length(unique(pcr_inconsistent$id[(!pcr_inconsistent$identical_race)]))
    )
  names(inconsistent_counts) <- c(
    "number_id_age_inconsistencies", 
    "number_id_sex_inconsistencies",
    "number_id_race_inconsistencies"
    )
  
  pcr_results_consistent <- pcr_results_adjusted[!(pcr_results_adjusted$id %in% pcr_inconsistent$id), ]

  
  # Need to keep all observations for a person up to and including their first positive
  first_pos <- pcr_results_consistent %>%
    filter(test_result == "positive") %>%
    group_by(id) %>%
    summarise(first_pos = min(posted_date))
  
  pcr_results_reduced <- left_join(pcr_results_consistent, first_pos) %>%
    mutate(first_pos = replace_na(first_pos, lubridate::ymd("9999-12-31"))) %>%
    filter(posted_date <= first_pos) %>%
    select(-first_pos) %>%
    distinct()
  
  
  # Add zip code level data and merge with pcr results
  zip_area_oc <- read_csv(
    here("data/zip-code-data", "zip-area2.csv"),
    col_types = cols(
      .default = col_skip(),
      NAME = col_character(),
      Zip = col_character(),
      AreaKm = col_double()
      )
    ) %>%
    select(name = NAME, zip = Zip, area_km = AreaKm)
  
  zip_pop_oc <- read_csv(
    here("data/zip-code-data", "zip-pop.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      Population = col_integer()
      )
    ) %>%
    drop_na() %>%
    mutate(population = Population / 1000) %>%
    select(zip = Zip, population)
  
  zip_data_merged <- merge(x = zip_area_oc, y = zip_pop_oc, by = "zip")
  zip_data_merged$pop_density <- zip_data_merged$population / zip_data_merged$area_km
  
  zip_income_oc <- read_csv(
    here("data/zip-code-data", "income-by-zip2.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      IncomeMed = col_integer(),
      IncPeriodofMeas = col_character()
      )
    ) %>%
    mutate(med_income = IncomeMed / 10000) %>%
    filter(IncPeriodofMeas == "2014-2018") %>%
    select(zip = Zip, med_income)
  
  zip_data_merged <- merge(x = zip_data_merged, y = zip_income_oc, by = "zip")

  zip_education_oc <- read_csv(
    here("data/zip-code-data", "education-by-zip.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      PercentBach = col_double()
      )
    ) %>%
    select(zip = Zip, percent_bachelors = PercentBach)
  
  zip_data_merged <- merge(x = zip_data_merged, y = zip_education_oc, by = "zip")

  zip_insurance_oc <- read_csv(
    here("data/zip-code-data", "insurance-by-zip.csv"),
    col_types = cols(
      .default = col_skip(),
      Zip = col_character(),
      PercentInsured = col_double()
      )
    ) %>%                           
    select(zip = Zip, percent_insured = PercentInsured)
  
  zip_data_merged <- merge(x = zip_data_merged, y = zip_insurance_oc, by = "zip")
  
  
  # Scale zip code level variables
  zip_data_merged$adj_pop_density <- scale(
    zip_data_merged$pop_density , 
    center = TRUE, 
    scale = TRUE
    )
  
  zip_data_merged$adj_med_income <- scale(
    zip_data_merged$med_income, 
    center = TRUE, 
    scale = TRUE
    )
  
  zip_data_merged$adj_perc_bach <- scale(
    zip_data_merged$percent_bachelors,
    center = TRUE,
    scale = TRUE
    )
  
  zip_data_merged$adj_perc_bach_quar <- with(
    zip_data_merged,
    cut(adj_perc_bach,
        breaks = quantile(adj_perc_bach, 
        probs = seq(0, 1, by = 0.25)),
        include.lowest = TRUE,
        labels = c("Q1", "Q2", "Q3", "Q4"))
    )
  
  zip_data_merged$adj_perc_insured <- scale(
    zip_data_merged$percent_insured,
    center = TRUE,
    scale = TRUE
    )
  
  zip_data_merged$adj_perc_insured_quar <- with(
    zip_data_merged,
    cut(
      adj_perc_insured,
      breaks = quantile(adj_perc_insured, probs = seq(0, 1, by = 0.25)),
      include.lowest = TRUE,
      labels = c("Q1", "Q2", "Q3", "Q4")
      )
    )
  
    #23 rows in 92678 zipcode we don't have area data for
  pcr_results_reduced$old_zip <- pcr_results_reduced$zip
  pcr_results_reduced$zip[pcr_results_reduced$old_zip == "92678"] <- "92679"
  pcr_results_merged <- merge(x = pcr_results_reduced, y = zip_data_merged, by = "zip")
  pcr_results_merged$old_zip <- factor(pcr_results_merged$old_zip)
  pcr_results_merged$zip <- factor(pcr_results_merged$zip)
  
  
  # Group ages 
  age_breaks <- c(0, 5, 10, 15, 20, 25, 30, 35, 40, 50, 60, 70, 80, 200)
  age_labels <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34","35-39",
                  "40-49","50-59","60-69","70-79","80+")
  setDT(pcr_results_merged)[, age_group := cut(
    age, 
    breaks = age_breaks, 
    right = FALSE, 
    labels = age_labels
    )]
  
  pcr_results_merged$age_group <- factor(
    pcr_results_merged$age_group,
    levels = age_labels
    )
  
  #technically we would want to do this differently for days but we don't used scaled in final model
   pcr_results_merged$adj_time_days <- scale( 
    pcr_results_merged$time_days,
    center = TRUE,
    scale = TRUE
    )
 
  # Count missingness by type
  missing_counts <- c(
    full_num_data_cases, 
    full_num_data_cases - nrow(pcr_results_adjusted), 
    nrow(pcr_results_adjusted) - nrow(pcr_results_consistent),
    nrow(pcr_results_consistent) - nrow(pcr_results_reduced),
    nrow(pcr_results_reduced) - nrow(pcr_results_merged)
    )
  missing_ids <- c(
    full_num_id_cases,
    full_num_id_cases - length(unique(pcr_results_adjusted$id)),
    length(unique(pcr_results_adjusted$id)) - length(unique(pcr_results_consistent$id)),
    length(unique(pcr_results_consistent$id)) - length(unique(pcr_results_reduced$id)),
    length(unique(pcr_results_reduced$id)) - length(unique(pcr_results_merged$id))
  )
  missing_tab <- data.frame("missing_counts" = missing_counts, "missing_ids" = missing_ids)
  rownames(missing_tab) <- c(
    "full_num_data_cases", 
    "num_na_cases_removed", 
    "num_inconsistent_cases_removed",
    "num_consecutive_cases_removed",
    "num_bad_zip_cases_removed"
  )

  
  list(
    "pcr_results_merged" = pcr_results_merged, 
    "zip_data_merged" = zip_data_merged, 
    "missing" = missing_tab,
    "inconsistencies" = inconsistent_counts
    )
}
```


```{r load-clean-data}
if (params$reclean_data) {
  all_pcr_and_zip <- read_all_pcr(
    file_path = here("data/positivity-data", "all-elr-pcr-tests-updated-2020-11-09.csv"),
    start_date = params$first_test_date,
    end_date = params$last_test_date
    )
  
  save(all_pcr_and_zip, file = here("data/positivity-data", "cleaned_process_pcr_data.Rdata"))
  
} 

load(file = here("data/positivity-data", "cleaned_process_pcr_data.Rdata"))
  
all_pcr <- data.frame(all_pcr_and_zip[["pcr_results_merged"]])
all_zip <- data.frame(all_pcr_and_zip[["zip_data_merged"]])
all_missing <- all_pcr_and_zip[["missing"]]
all_inconsist_ids <- all_pcr_and_zip[["inconsistencies"]]


# new data added after 1st review
house_data_og <- read_csv(here("data/zip-code-data", "house-crowding.csv")) %>% 
  janitor::clean_names()

all_pcr <- house_data_og %>% 
  filter(period_of_measure == "2015-2019") %>% 
  filter(breakout_subcategory == "Owners") %>% 
  select(zip = location, house_crowding = indicator_rate_value) %>% 
  mutate(zip = factor(zip)) %>% 
  right_join(y = all_pcr, by = "zip")


```


```{r missing-data-table}
usable_counts <- c(
  all_missing[1, "missing_counts"],
  all_missing[1, "missing_counts"] - all_missing[2, "missing_counts"],
  all_missing[1, "missing_counts"] - sum(all_missing[2:3, "missing_counts"]),
  all_missing[1, "missing_counts"] - sum(all_missing[2:4, "missing_counts"]),
  all_missing[1, "missing_counts"] - sum(all_missing[2:5, "missing_counts"])
)
usable_ids <- c(
  all_missing[1, "missing_ids"],
  all_missing[1, "missing_ids"] - all_missing[2, "missing_ids"],
  all_missing[1, "missing_ids"] - sum(all_missing[2:3, "missing_ids"]),
  all_missing[1, "missing_ids"] - sum(all_missing[2:4, "missing_ids"]),
  all_missing[1, "missing_ids"] - sum(all_missing[2:5, "missing_ids"])
)

tab_obs <- data.frame(
  c(NA, all_missing[2:5, "missing_counts"]), 
  usable_counts, 
  c(NA, all_missing[2:5, "missing_ids"]), 
  usable_ids
)
rownames(tab_obs) <- c(
  "Unmodified",
  "Missing or Inconclusive[note]",
  "Inconsistent",
  "Consecutive Testing[note]",
  "Invalid Zip Code"
)

kable(
  tab_obs,
  col.names = c(
    "With Issue", 
    "After Removal",
    "With Issue ", 
    "After Removal "
  ),
  format = "latex",
  caption = paste(
    "Number of SARS-CoV-2 tests and unique individuals who tested in Orange County from ",
      params$first_test_date, " to ", params$last_test_date,
      ". Includes number removed from data by type of issue for this analysis.",
    sep = "")
  ) %>%
  kable_styling(latex_options = c("HOLD_position")) %>% 
  add_header_above(c(" " = 1, "Number of Tests" = 2, "Number of Indiviuals" = 2), line = FALSE) %>% 
  add_footnote(
    c(
      "Number without conclusive test result or complete demographic information",
      "Kept an indivdual's tests up to and including their first positive"
      ),
    notation = "symbol"
    )
```


```{r compute-ci-gam-logistic-function}
compute_ci_gam_logistic <- function(model, alpha = 0.05, param_names) {
  #does not include intercept or median income
  model_summary <- summary(model)
  in_med_inc <- length(model_summary$p.coeff)
  coeffs <- model_summary$p.coeff[-c(1, in_med_inc)]
  se <- sqrt(diag(vcov(model, unconditional = TRUE)))[2:(length(coeffs) + 1)]
  model_sum <- data.frame("odds" = exp(coeffs),
                          "lower_bound" = exp(coeffs - qnorm(1 - alpha / 2) * se),
                          "upper_bound" = exp(coeffs + qnorm(1 - alpha / 2) * se),
                          "p-value" =  model_summary$p.pv[-c(1, in_med_inc)],
                          "param_names" = param_names)
  
  return("model_sum" = model_sum)
}
```

For simplicity let  
$O_i$ be the odds of testing positive for SARS-CoV-2 in Orange County.  
$\overrightarrow{\beta}_{\text{Age Group}} = (\beta_{\text{Age[5-9]}}, \beta_{\text{Age[10-14]}}, \beta_{\text{Age[15-19]}}, \beta_{\text{Age[20-24]}}, \beta_{\text{Age[25-29]}}, \beta_{\text{Age[30-34]}}, \beta_{\text{Age[35-39]}}\\ \beta_{\text{Age[40-49]}},  \beta_{\text{Age[50-59]}}, \beta_{\text{Age[60-69]}}, \beta_{\text{Age[70-79]}}, \beta_{\text{Age[80+]}})$  
$\overrightarrow{\beta}_{\text{Race}} = (\beta_{\text{Asian}}, \beta_{\text{Black}}, \beta_{\text{Hispanic}}, \beta_{\text{NativeAmerican}}, \beta_{\text{PacificIslander}}, \beta_{\text{OtherRace}}, \beta_{\text{Unknown}})$  
$\overrightarrow{\beta}_{\text{College}} = (\beta_{\text{\% with College Degree Quartile 2}}, \beta_{\text{\% with College Degree Quartile 3}}, \beta_{\text{\% with College Degree Quartile 4}})$  
$\overrightarrow{\beta}_{\text{Insurance}} = (\beta_{\text{\% with Insurance Quartile 2}}, \beta_{\text{\% with Insurance Quartile 3}}, \beta_{\text{\% with Insurance Quartile 4}})$


Model 1:
\begin{equation}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}\text{Time}_i,
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model1}
if (params$refit_model1) {
  tic()
  fit_time_lin1 <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days +
              (1 | zip),              
    family = binomial, 
    data = all_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  skrrrahh("drummaboy")
  # 4423.84 sec elapsed
  # Model failed to converge with max|grad| = 0.00298652 (tol = 0.002, component 1)
  tic()
  ss <- getME(fit_time_lin1, c("theta","fixef"))
  fit_time_lin2 <- update(
    fit_time_lin1,
    start = ss,
    control = glmerControl(
      optimizer = "bobyqa",
      optCtrl = list(maxfun = 2e6)
    )
  )
  toc()
  fit_time_lin <- fit_time_lin2
  skrrrahh("drummaboy")
  # 4372.56 sec elapsed
  # updated and saved 11-3-2020
  save(fit_time_lin, file = here("analysis/testing-positive-regression-results", "fit_time_lin.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_lin.Rdata"))
}
```

Model 2:
\begin{equation}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}\text{Time}_i + \beta_{\text{Time2}}\text{Time}^2_i,
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model2}
if (params$refit_model2) {
  tic()
  fit_time_quad <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              I(adj_time_days) + I(adj_time_days^2) + 
              (1 | zip),              
    family = binomial, 
    data = all_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  skrrrahh("drummaboy")
  # 4846.83 sec elapsed
  # updated and saved 11-2-2020
  save(fit_time_quad, file = here("analysis/testing-positive-regression-results", "fit_time_quad.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_quad.Rdata"))
}
```

Model 3:
\begin{equation}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}\text{Time}_i + \beta_{\text{Time2}}\text{Time}^2_i, \\
		&+ \beta_{\text{Interaction1}}\text{Median Income}_i\times\text{Time}_i + \beta_{\text{Interaction2}}\text{Median Income}_i\times\text{Time}^2_i,
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model3}
if (params$refit_model3) {
  tic()
  fit_time_quad_inter1 <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days + I(adj_time_days^2) + 
              adj_time_days:adj_med_income + I(adj_time_days^2):adj_med_income +
              (1 | zip),              
    family = binomial, 
    data = all_pcr,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  # 5212.67 sec elapsed
  # Model failed to converge with max|grad| = 0.00348316 (tol = 0.002, component 1)
    tic()
  ss <- getME(fit_time_quad_inter1, c("theta","fixef"))
  fit_time_quad_inter2 <- update(
    fit_time_quad_inter1,
    start = ss,
    control = glmerControl(
      optimizer = "bobyqa",
      optCtrl = list(maxfun = 2e6)
    )
  )
  toc()

  fit_time_quad_inter <- fit_time_quad_inter2
  # 3284.9 sec elapsed
  # Model failed to converge with max|grad| = 0.00348316 (tol = 0.002, component 1)
  skrrrahh("drummaboy")
  # updated 11/3/2020
  save(fit_time_quad_inter, file = here("analysis/testing-positive-regression-results", "fit_time_quad_inter.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_quad_inter.Rdata"))
}
```

Model 4:  
Generalized additive model
\begin{equation}
	\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}f(\text{Time}_i),
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model4}
if (params$refit_model4) {
  tic()
  fit_time_gam <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density  + adj_med_income +
              s(time_days, bs = "ts", k = -1) + 
              s(zip, bs = "re"),              
    family = binomial, 
    data = all_pcr,
    method = "REML",
    gamma = 1.5
    )
  toc()
  skrrrahh("drummaboy")
  # 1289.41 sec elapsed
  # updated and saved 11-2-2020
  save(fit_time_gam, file = here("analysis/testing-positive-regression-results", "fit_time_gam.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_gam.Rdata"))
}
```

Model 5:  
Generalized additive model
\begin{equation}
		\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}f(\text{Time}_i) + \beta_{\text{Interaction}}\text{Median Income}_i\times f(\text{Time}_i),
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model5, results="hide"}
if (params$refit_model5) {
  tic()
  fit_time_gam_inter <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density  + adj_med_income +
              s(time_days) +
              ti(time_days, adj_med_income, bs = "ts") + 
              s(zip, bs = "re"),              
    family = binomial, 
    data = all_pcr,
    method = "REML",
    gamma = 1.5
    )
  toc()
  skrrrahh("drummaboy")
# 4714.62 sec elapsed
# updated and saved 11-2-2020
  save(fit_time_gam_inter, file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter.Rdata"))
}
 
parameter_names <- c(
  "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", 
  "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Male",
  "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Other", "Unknown",
  "2nd Quartile", "3rd Quartile", "4th Quartile",
  " 2nd Quartile", " 3rd Quartile", " 4th Quartile",
  "Population Density"
  )

fit_time_gam_inter_summary1 <- compute_ci_gam_logistic(
  model = fit_time_gam_inter,
  param_names = parameter_names
)

gam_smooth_var_ci <- gam.vcomp(fit_time_gam_inter)
```


Model 6:  
Generalized additive model
\begin{equation}
		\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}f(\text{Time}_i) + \beta_{\text{Interaction}}\text{Median Income}_i\times f(\text{Time}_i),
	\end{split}
\end{equation}
without a random intercept for zip code.

```{r fit-model6, results="hide"}
if (params$refit_model6) {
  tic()
  fit6 <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density  + adj_med_income +
              s(time_days) +
              ti(time_days, adj_med_income, bs = "ts"),
    family = binomial, 
    data = all_pcr,
    method = "REML",
    gamma = 1.5
    )
  toc()
  skrrrahh("drummaboy")
# 4714.62 sec elapsed
# updated and saved 11-2-2020
  save(fit6, file = here("analysis/testing-positive-regression-results", "fit6.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit6.Rdata"))
}
```


Model 7:  
Generalized additive model
\begin{equation}
		\begin{split}
		\log(O_i) &= \beta_0 + \overrightarrow{\beta}_{\text{Age Group}}\overrightarrow{\text{Age Group}}_i + \beta_{\text{Sex}}\text{Sex}_i + \overrightarrow{\beta}_{\text{Race}}\overrightarrow{\text{Race}}_i\\
		&+ \overrightarrow{\beta}_{\text{College}}\overrightarrow{\text{\% with College Degree Quartile}}_i\\
		&+ \overrightarrow{\beta}_{\text{Insurance}}\overrightarrow{\text{\% with Medical Insurance Quartile}}_i \\
		&+ \beta_{\text{Population Density}}\text{Population Density}_i + \beta_{\text{House Crowding}}\text{House Crowding}_i + \beta_{\text{Median Income}}\text{Median Income}_i\\ 
		&+ \beta_{\text{Time}}f(\text{Time}_i) + \beta_{\text{Interaction}}\text{Median Income}_i\times f(\text{Time}_i),
	\end{split}
\end{equation}
with a random intercept for zip code.

```{r fit-model7, results="hide"}
if (params$refit_model7) {
  tic()
  fit_time_gam_inter_house <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + house_crowding + adj_med_income +
              s(time_days) +
              ti(time_days, adj_med_income, bs = "ts") + 
              s(zip, bs = "re"),              
    family = binomial, 
    data = all_pcr,
    method = "REML",
    gamma = 1.5
    )
  toc()
  skrrrahh("drummaboy")
# 4952.92 sec elapsed
# updated and saved 2021-05-24
  save(fit_time_gam_inter_house, file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter_house.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter_house.Rdata"))
}
 
parameter_names <- c(
  "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", 
  "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Male",
  "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Other", "Unknown",
  "2nd Quartile", "3rd Quartile", "4th Quartile",
  " 2nd Quartile", " 3rd Quartile", " 4th Quartile",
  "Population Density",
  "House Crowding"
  )

fit_time_gam_inter_house_summary1 <- compute_ci_gam_logistic(
  model = fit_time_gam_inter_house,
  param_names = parameter_names
)

gam_smooth_var_ci <- gam.vcomp(fit_time_gam_inter_house)
```


```{r bic-table}
compare_bic <- BIC(fit_time_lin, fit_time_quad, fit_time_quad_inter, fit_time_gam, fit_time_gam_inter, fit6, fit_time_gam_inter_house) 
row.names(compare_bic) <- c("Model 1", "Model 2", "Model 3", "Model 4", "Model 5", "Model 6", "Model 7")

kable(compare_bic,
      col.names = c("Degrees of Freedom", "BIC"),
      format = "latex",
      caption = "Model comparison using BIC supports Model 5 to model odds of testing positive for COVID-19 in Orange County.") %>%
  kable_styling(latex_options = "HOLD_position")
```


```{r final-model-results-table}
fit_final_model <- fit_time_gam_inter_summary1

rr_fm <- round(fit_final_model[, "odds"], 3)
ci_lb_fm <- round(fit_final_model[, "lower_bound"], 2)
ci_ub_fm <- round(fit_final_model[, "upper_bound"], 2)
#p_values_fm <- fit_final_model$p.value
#p_values_fm <- ifelse(p_values_fm <= 0.001, "0.000", paste(round(p_values_fm, 3)))
# The 5 accounts for the reference groups
table_fm <- data.frame(matrix(NA, nrow = nrow(fit_final_model) + 5, ncol = 3))
colnames(table_fm) <- c("SARS-CoV-2+", "Total", "with (95% CI[note])")
rownames(table_fm) <- c(
  "0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", 
  "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Female", "Male",
  "White", "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Other", "Unknown",
  "1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile",
  "1st Quartile ", "2nd Quartile ", "3rd Quartile ", "4th Quartile ",
  "Population Density (1000ppl/km^2)"
  )
reference_group <- c(
  TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,
  TRUE, FALSE,
  TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  FALSE
  )
          
non_ref_counter <- 1
for(i in 1:nrow(table_fm)) {
  if (reference_group[i]) {
    #table_fm[i, 1] <- "Reference"
    table_fm[i, 3] <- "Reference"
  } else {
    table_fm[i, 3] <- paste(rr_fm[non_ref_counter], " (", 
                          ci_lb_fm[non_ref_counter], ", ", 
                          ci_ub_fm[non_ref_counter], ")", 
                          sep = "")
    #table_fm[i, 2] <- p_values_fm[non_ref_counter]
    non_ref_counter <- non_ref_counter + 1
  }
}      

variable_counts <- data.frame(matrix(NA,  nrow = nrow(fit_final_model) + 5, ncol = 2))
all_pcr_ppl <- all_pcr[!duplicated(all_pcr$id), ]
cov_pos_pcr_ppl <- all_pcr_ppl[all_pcr_ppl$covid_positive == 1, ]
quant_vars <- c("pop_density")
order_quant_vars <- c(nrow(table_fm))
variable_counts[order_quant_vars, c(1, 2)] <- c("", "")

qual_vars <- c("age_group", "sex", "race", "adj_perc_bach_quar", "adj_perc_insured_quar")
order_qual_vars <- (1:nrow(table_fm))[-order_quant_vars]

qual_vars_ordered <- c(
  "age_group.0-4","age_group.5-9", "age_group.10-14", "age_group.15-19",
  "age_group.20-24", "age_group.25-29", "age_group.30-34", "age_group.35-39",
  "age_group.40-49", "age_group.50-59", "age_group.60-69", "age_group.70-79",
  "age_group.80+", 
  "sex.female", "sex.male", 
  "race.white", "race.asian",
  "race.black or african american", "race.hispanic or latino",
  "race.american indian or alaska native", "race.native hawaiian or other pacific islander",
  "race.other", "race.unknown",
  "adj_perc_bach_quar.Q1", "adj_perc_bach_quar.Q2",
  "adj_perc_bach_quar.Q3", "adj_perc_bach_quar.Q4",
  "adj_perc_insured_quar.Q1", "adj_perc_insured_quar.Q2",
  "adj_perc_insured_quar.Q3", "adj_perc_insured_quar.Q4"
  )
 
calc_freq <- function(df, order_vars) {
  qual_count <- unlist(apply(df, MARGIN = 2, FUN = table))

  qual_count <- qual_count[order_vars]
  qual_freq <- round(100 * qual_count / nrow(df), 2)
  paste(qual_count, " (", qual_freq, "%)", sep = "")
}

variable_counts[order_qual_vars, 2] <- calc_freq(all_pcr_ppl[, qual_vars], qual_vars_ordered)
variable_counts[order_qual_vars, 1] <- calc_freq(cov_pos_pcr_ppl[, qual_vars], qual_vars_ordered)

table_fm[, 2] <- variable_counts[, 2]
table_fm[, 1] <- variable_counts[, 1]

fm_sum <- summary(fit_time_gam)
intercept <- paste0(
  round(exp(fm_sum$p.coeff[1]), 3),
  " (",
  round(exp(fm_sum$p.coeff[1] - qnorm(0.975) * fm_sum$se[1]), 2),
  ",",
  round(exp(fm_sum$p.coeff[1] + qnorm(0.975) * fm_sum$se[1]), 2),
  ")"
)


kable(
  table_fm, 
  format = "latex", 
  caption = paste(
    "Model 5 logistic regression results for linear fixed effects of odds of testing positive for SARS-CoV-2 in Orange County. This table excludes the coefficient for median income due to the interaction between median income and time.", 
    sep = ""
    )
  ) %>%
  kable_styling(latex_options = c("HOLD_position"), font_size = 11) %>%
  pack_rows("Age", 1, 13, bold = FALSE) %>%
  pack_rows("Sex", 14,15, bold = FALSE) %>%
  pack_rows("Race", 16, 23, bold = FALSE) %>%
  pack_rows("% with College Degree[note]", 24, 27, bold = FALSE) %>%
  pack_rows("% with Insurance", 28, 31, bold = FALSE) %>% 
  add_header_above(c(" " = 1, "Counts" = 2, "Adjusted Odds Ratio[note]" = 1), line = FALSE) %>% 
  add_footnote(
    c(
      paste0(
        "Adjusted for all covariates listed plus zip code estimated median income and time of test in days. Model intercept represents odds of a white female in the 0 to 4 age group in a zip code in the first quartile of college degree and insured with the average population density. The odds of this individual testing positive for COVID-19 is estimated to be ",
        intercept
      ), 
      "95% Confidence Interval",
      "Esimated: percent of people with a bachelor's degree, percent of people with medical insurance, and population density in an individual's zip code "
      ),
    notation = "symbol"
  )  
```


```{r final-model-results-coeff-plot, fig.align="center", fig.cap="Model 5 logistic regression results for linear fixed effects of odds of testing positive for SARS-CoV-2 in Orange County. This plot excludes the coefficient for median income due to the interaction between median income and time. Percent with college degree is the estimated associated change in odds of testing positive for SARS-CoV-2 for an individual living in a zip code where the estimate percent of the population with a bachelor's degree is in a given quartile relative to an individual from a zip code in the first. Percent insured representing estimated percent of zip code population with medical insurance is similarly defined."}

fm_results_with_ref <- rbind(
  c(1, 0.99, 1.01, 0, paste0(c("Age", rep(" ", 34), "[ref: 0-4 years]"), collapse = "")),
  fit_time_gam_inter_summary1[1:12, ],
  c(1, 0.99, 1.01, 0, paste0(c("Sex", rep(" ", 38), "[ref: Female]"), collapse = "")),
  fit_time_gam_inter_summary1[13, ],
  c(1, 0.99, 1.01, 0, paste0(c("Race/Ethnicity", rep(" ", 24), "[ref: White]"), collapse = "")),
  fit_time_gam_inter_summary1[14:20, ],
  c(1, 0.99, 1.01, 0, "% with College Degree [ref: 1st Quartile]"),
  fit_time_gam_inter_summary1[21:23, ],
  c(1, 0.99, 1.01, 0, paste0(c("% with Insurance", rep(" ", 10), "[ref: 1st Quartile]"), collapse = "")),
  fit_time_gam_inter_summary1[24:26, ],
  c(as.character(fit_time_gam_inter_summary1[27, -5]), "Population Density (1000ppl/km^2)")
)

fm_results_with_ref$odds <- as.double(fm_results_with_ref$odds)
fm_results_with_ref$lower_bound <- as.double(fm_results_with_ref$lower_bound)
fm_results_with_ref$upper_bound <- as.double(fm_results_with_ref$upper_bound)

variable_colors <- c(
  "white",
  rep("black", 12),
  "white",
  rep("black", 1),
  "white",
  rep("black", 7),
  "white",
  rep("black", 3),
  "white",
  rep("black", 4)
)


ggplot(fm_results_with_ref, aes(x = param_names, y = odds), size = 5) +
  ylim(min(fm_results_with_ref$lower_bound), max(fm_results_with_ref$upper_bound)) +
  geom_point(color = variable_colors) +
  geom_linerange(aes(ymin = lower_bound, ymax = upper_bound), color = variable_colors) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
  ylab("Adjusted Odds Ratio") +
  coord_flip() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x=element_text(size = 12, colour="black"),
    axis.text.y=element_text(colour="black", hjust = 1),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "black"),
    panel.grid.major = element_line(colour = "gray"),
    panel.background = element_blank(),
    plot.title = element_text(hjust=0)
  ) +
  scale_x_discrete(limits = rev(fm_results_with_ref$param_names)) +
  scale_y_continuous(breaks = seq(0, 2.5, by = 0.5)) +
  ggtitle("Generalized additive logistic regression results for odds ratio of testing SARS-CoV-2 positive \n in Orange County")
```


```{r final-model-results-interaction-plot, fig.align="center", fig.width=7, fig.height=7, fig.cap="Three dimentional visualization of interaction between time and median income from logistic model 5. SARS-CoV-2 test results in Orange County from March 1st to August 16th. Large difference visible in early March between individuals in zip codes with relatively high median incomes and those in zip codes with relatively small median incomes; this difference diminishes as time increase. Gap is due to lack of zip codes with median income in interval, the extreme median income is Newport Coast with zip code 92657."}
plot(fit_time_gam_inter, scheme = 1, select = 2, 
     xlab = "Time (std. dev.)",
     ylab = "Median Income (std. dev.)",
     main = "Interaction") 
```

