---
title: "Positivity Modeling"
author: "Christie Yang"
date: "5/20/2022"
output: html_document

params:
  reclean_data: FALSE
  refit_model1: FALSE
  refit_model2: FALSE
  refit_model3: FALSE
  refit_model4: FALSE
  refit_model5: FALSE

---

## Positivity Data Modeling

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(dplyr)
library(here)
library(tictoc)     
library(lme4)       # GLM
library(mgcv)       # GAM modeling
library(kableExtra) # table formatting
library(mgcv.helper)

# Positivity Data
load(here("data/cleaned-data", "usable_tests.Rdata"))

```

- All of the following models have a random intercept for ZIP code.
- Population density to replace house crowding density.

### Model 1 --------------------------------------------------------------------

- Time is linear (untransformed)
- No interaction term between median household income and time
```{r}
if (params$refit_model1) {
  tic()
  fit_time_lin <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days +
              (1 | zip),              
    family = binomial, 
    data = usable_tests,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  # 
  # 2730.94 sec elapsed
  # updated and saved 2021-06-04
  save(fit_time_lin, file = here("analysis/testing-positive-regression-results", "fit_time_lin.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_lin.Rdata"))
}
```


### Model 2 --------------------------------------------------------------------

- Time is quadratic to see if the rate of change in the mean response depends on time
- No interaction term between median household income and time
```{r}
if (params$refit_model2) {
  tic()
  fit_time_quad <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              I(adj_time_days) + I(adj_time_days^2) + 
              (1 | zip),              
    family = binomial, 
    data = usable_tests,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  # 
  # 2916.06 sec elapsed
  # Updated 2020-06-04
  save(fit_time_quad, file = here("analysis/testing-positive-regression-results", "fit_time_quad.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_quad.Rdata"))
}
```


### Model 3 --------------------------------------------------------------------

- time is quadratic to see if the rate of change in the mean response depends on time
- interaction term between median household income and time
```{r}
if (params$refit_model3) {
  tic()
  fit_time_quad_inter <- glmer(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density + adj_med_income +
              adj_time_days + I(adj_time_days^2) + 
              adj_time_days:adj_med_income + I(adj_time_days^2):adj_med_income +
              (1 | zip),              
    family = binomial, 
    data = usable_tests,
    control = glmerControl(optimizer ="bobyqa", optCtrl = list(maxfun = 2e6))
    )
  toc()
  # 2903.28 sec elapsed
  # updated 2021-06-04
  save(fit_time_quad_inter, file = here("analysis/testing-positive-regression-results", "fit_time_quad_inter.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_quad_inter.Rdata"))
}
```


### Model 4 --------------------------------------------------------------------

- GAM: used when the linear predictor depends on __unknown__ smoothed functions
- GAM preferred over GLM because we want to replace the linear term (time variable) with a non-linear function. 
- Specified bs = "ts" in smoothing function to indicate the type of spline
- No interaction term between median household income and time

```{r}
if (params$refit_model4) {
  tic()
  fit_time_gam <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density  + adj_med_income +
              s(time_days, bs ="ts") + 
              s(zip, bs = "re"), # random intercept in GAM             
    family = binomial, 
    data = usable_tests,
    method = "REML",
    gamma = 1.5
    )
  toc()
  # 
  # updated and saved 2021-06-04
  # 915.72 sec elapsed
  save(fit_time_gam, file = here("analysis/testing-positive-regression-results", "fit_time_gam.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_gam.Rdata"))
}
```

### Model 5 --------------------------------------------------------------------

- Identical to Model 4 but added back interaction term between median household
income and time

```{r}
if (params$refit_model5) {
  tic()
  fit_time_gam_inter <- gam(
    formula = covid_positive ~ age_group + sex + race + 
              adj_perc_bach_quar + adj_perc_insured_quar +
              adj_pop_density  + adj_med_income +
              s(time_days) +
              ti(time_days, adj_med_income, bs = "ts") + 
              s(zip, bs = "re"),              
    family = binomial, 
    data = usable_tests,
    method = "REML",
    gamma = 1.5
    )
  toc()
  # 
# 1889.62 sec elapsed
# updated and saved 2021-06-04
  save(fit_time_gam_inter, file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter.Rdata"))
} else {
  load(file = here("analysis/testing-positive-regression-results", "fit_time_gam_inter.Rdata"))
}

summary(fit_time_gam_inter)

```


# BIC Scores

```{r}
# Model 5 had the lowest BIC score 
compare_bic <- BIC(fit_time_lin, fit_time_quad, fit_time_quad_inter, fit_time_gam, fit_time_gam_inter) 
row.names(compare_bic) <- c("Model 1", "Model 2", "Model 3", "Model 4", "Model 5")
compare_bic %>% kbl(caption = "Model Comparison using BIC") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

```{r}
##### function 
compute_ci_gam_logistic <- function(model, alpha = 0.05, param_names) {
  #does not include intercept or median income
  model_summary <- summary(model)
  in_med_inc <- length(model_summary$p.coeff)
  coeffs <- model_summary$p.coeff[-c(1, in_med_inc)]
  se <- sqrt(diag(vcov(model, unconditional = TRUE)))[2:(length(coeffs) + 1)]
  model_sum <- data.frame(
    "odds" = exp(coeffs),
    "lower_bound" = exp(coeffs - qnorm(1 - alpha / 2) * se),
    "upper_bound" = exp(coeffs + qnorm(1 - alpha / 2) * se),
    "p-value" =  model_summary$p.pv[-c(1, in_med_inc)],
    "param_names" = param_names
  )
  
  return("model_sum" = model_sum)
}

parameter_names <- c(
  "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", 
  "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Male",
  "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Unknown",
  "2nd Quartile", "3rd Quartile", "4th Quartile",
  " 2nd Quartile", " 3rd Quartile", " 4th Quartile",
  "Population Density"
  )

model5 <- compute_ci_gam_logistic(
  model = fit_time_gam_inter,
  param_names = parameter_names
)

gam_smooth_var_ci <- gam.vcomp(fit_time_gam_inter)
#####
```

```{r final-model-results-table, , results='asis'}
fit_final_model <- model5

rr_fm <- round(fit_final_model[, "odds"], 3)
ci_lb_fm <- round(fit_final_model[, "lower_bound"], 2)
ci_ub_fm <- round(fit_final_model[, "upper_bound"], 2)
#p_values_fm <- fit_final_model$p.value
#p_values_fm <- ifelse(p_values_fm <= 0.001, "0.000", paste(round(p_values_fm, 3)))
# The 5 accounts for the reference groups
table_fm <- data.frame(matrix(NA, nrow = nrow(fit_final_model) + 5, ncol = 3))
colnames(table_fm) <- c("SARS-CoV-2+", "Total", "with (95% CI[note])")
rownames(table_fm) <- c(
  "0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", 
  "35-39", "40-49", "50-59", "60-69", "70-79", "80+",
  "Female", "Male",
  "White", "Asian", "Black", "Hispanic", "Native American", "Pacific Islander", "Unknown",
  "1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile",
  "1st Quartile ", "2nd Quartile ", "3rd Quartile ", "4th Quartile ",
  "Population Density (1000ppl/km^2)"
  )
reference_group <- c(
  TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,
  TRUE, FALSE,
  TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  TRUE, FALSE, FALSE, FALSE,
  FALSE
  )
          
non_ref_counter <- 1
for(i in 1:nrow(table_fm)) {
  if (reference_group[i]) {
    #table_fm[i, 1] <- "Reference"
    table_fm[i, 3] <- "Reference"
  } else {
    table_fm[i, 3] <- paste(rr_fm[non_ref_counter], " (", 
                          ci_lb_fm[non_ref_counter], ", ", 
                          ci_ub_fm[non_ref_counter], ")", 
                          sep = "")
    #table_fm[i, 2] <- p_values_fm[non_ref_counter]
    non_ref_counter <- non_ref_counter + 1
  }
}      

variable_counts <- data.frame(matrix(NA,  nrow = nrow(fit_final_model) + 4, ncol = 2))
all_pcr_ppl <- usable_tests[!duplicated(usable_tests$id), ]
cov_pos_pcr_ppl <- all_pcr_ppl[all_pcr_ppl$covid_positive == 1, ]
quant_vars <- c("pop_density")
order_quant_vars <- c(nrow(table_fm) - 1)
variable_counts[order_quant_vars, c(1)] <- c("")

qual_vars <- c("age_group", "sex", "race", "adj_perc_bach_quar", "adj_perc_insured_quar")
order_qual_vars <- (1:nrow(table_fm))[-order_quant_vars]

qual_vars_ordered <- c(
  "age_group.0-4","age_group.5-9", "age_group.10-14", "age_group.15-19",
  "age_group.20-24", "age_group.25-29", "age_group.30-34", "age_group.35-39",
  "age_group.40-49", "age_group.50-59", "age_group.60-69", "age_group.70-79",
  "age_group.80+", 
  "sex.female", "sex.male", 
  "race.white", "race.asian",
  "race.black", "race.hispanic",
  "race.native", "race.islander",
  "race.unknown",
  "adj_perc_bach_quar.Q1", "adj_perc_bach_quar.Q2",
  "adj_perc_bach_quar.Q3", "adj_perc_bach_quar.Q4",
  "adj_perc_insured_quar.Q1", "adj_perc_insured_quar.Q2",
  "adj_perc_insured_quar.Q3", "adj_perc_insured_quar.Q4"
  )
 
calc_freq <- function(df, order_vars) {
  qual_count <- unlist(apply(df, MARGIN = 2, FUN = table))

  qual_count <- qual_count[order_vars]
  qual_freq <- round(100 * qual_count / nrow(df), 2)
  paste(qual_count, " (", qual_freq, "%)", sep = "")
}

variable_counts[order_qual_vars, 2] <- calc_freq(all_pcr_ppl[, qual_vars], qual_vars_ordered)
variable_counts[order_qual_vars, 1] <- calc_freq(cov_pos_pcr_ppl[, qual_vars], qual_vars_ordered)

table_fm[, 2] <- variable_counts[, 2]
table_fm[, 1] <- variable_counts[, 1]

fm_sum <- summary(fit_time_gam)
intercept <- paste0(
  round(exp(fm_sum$p.coeff[1]), 3),
  " (",
  round(exp(fm_sum$p.coeff[1] - qnorm(0.975) * fm_sum$se[1]), 2),
  ",",
  round(exp(fm_sum$p.coeff[1] + qnorm(0.975) * fm_sum$se[1]), 2),
  ")"
)

model5 %>% kbl(caption = "Model 5 Logistic Results ") %>%
  kable_classic(full_width = F, html_font = "Cambria")

table_fm %>% kbl(caption =  "Model 5 logistic regression results for linear fixed effects of odds of testing positive for SARS-CoV-2 in Orange County. This table excludes the coefficient for median income due to the interaction between median income and time.") %>% 
  kable_classic(full_width = F) %>% 
  pack_rows("Age", 1, 13, bold = FALSE) %>%
  pack_rows("Gender", 14,15, bold = FALSE) %>%
  pack_rows("Race/ethnicity", 16, 23, bold = FALSE) %>%
  pack_rows("% with College Degree[note]", 24, 27, bold = FALSE) %>%
  pack_rows("% with Insurance", 28, 31, bold = FALSE) %>% 
  add_header_above(c(" " = 1, "Counts" = 2, "Adjusted Odds Ratio[note]" = 1), line = FALSE)

```

```{r fig.align="center", fig.cap="Figure 5. Model 5 results for each fixed effects odds of testing positive for COVID-19 in Orange County. Note that median income and time were excluded due to the interaction term. Groups 10-14 and 15-19 had almost 1.8 times the odds of testing positive than individuals in groups 0-4. A dotted line was also added to help illustrate which features were significant."}

fm_results_with_ref <- rbind(
  c(1, 0.99, 1.01, 0, paste0(c("Age", rep(" ", 34), "[ref: 0-4 years]"), collapse = "")),
  model5[1:12, ],
  c(1, 0.99, 1.01, 0, paste0(c("Gender", rep(" ", 33), "[ref: Female]"), collapse = "")),
  model5[13, ],
  c(1, 0.99, 1.01, 0, paste0(c("Race/ethnicity", rep(" ", 25), "[ref: White]"), collapse = "")),
  model5[14:19, ],
  c(1, 0.99, 1.01, 0, "% with College Degree [ref: 1st Quartile]"),
  model5[20:22, ],
  c(1, 0.99, 1.01, 0, paste0(c("% with Insurance", rep(" ", 10), "[ref: 1st Quartile]"), collapse = "")),
  model5[23:25, ],
  c(as.character(model5[26, -5]), "Population Density (1000ppl/km^2)")
)

fm_results_with_ref$odds <- as.double(fm_results_with_ref$odds)
fm_results_with_ref$lower_bound <- as.double(fm_results_with_ref$lower_bound)
fm_results_with_ref$upper_bound <- as.double(fm_results_with_ref$upper_bound)

variable_colors <- c(
  "white",
  rep("black", 12),
  "white",
  rep("black", 1),
  "white",
  rep("black", 6),
  "white",
  rep("black", 3),
  "white",
  rep("black", 4)
)


ggplot(fm_results_with_ref, aes(x = param_names, y = odds), size = 5) +
  ylim(min(fm_results_with_ref$lower_bound), max(fm_results_with_ref$upper_bound)) +
  geom_point(color = variable_colors) +
  geom_linerange(aes(ymin = lower_bound, ymax = upper_bound), color = variable_colors) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
  ylab("Adjusted Odds Ratio") +
  coord_flip() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x=element_text(size = 12, colour="black"),
    axis.text.y=element_text(colour="black", hjust = 1),
    axis.ticks.y = element_blank(),
    axis.line.x = element_line(color = "black"),
    panel.grid.major = element_line(colour = "gray"),
    panel.background = element_blank(),
    plot.title = element_text(hjust=0),
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0)
  ) + 
  scale_x_discrete(limits = rev(fm_results_with_ref$param_names)) +
  scale_y_continuous(breaks = seq(0, 2.5, by = 0.5)) +
  ggtitle("Logistic regression odds ratio of testing COVID positive \n in Orange County")

```





